---
title: 'Status of Yelloweye Rockfish off the U.S. West Coast in 2025'
author:
# add authors in order
  - name: 'Morgan A. Johnston*'
    affiliations:
      - name: 'Oregon State University'
        address: '1500 SW Jefferson Way'
        city: 'Corvallis, OR'
        state: 'OR'
        postal-code: '97331'
  - name: 'R. Claire Rosemond*'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2032 SE OSU Drive Building 955' 
        city: 'Newport, OR' 
        state: 'OR' 
        postal-code: '98112-2097'
  - name: 'Alison Whitman'
    affiliations:
      - name: 'Oregon Department of Fish and Wildlife'
        address: '2040 SE Marine Science Drive'
        city: 'Newport, OR'
        state: 'OR'
        postal-code: '97365'
  - name: 'Elizabeth Perl'
    affiliations:
      - name: 'ECS Federal in support of NMFS OST'
        address: 'East-West Hwy'
        city: 'Silver Spring, MD'
        state: 'MD'
        postal-code: '22031'
  - name: 'Matheus de Barros'
    affiliations:
      - name: 'University of Washington'
        address: '1410 NE Campus Pkwy'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98195'
  - name: 'Juliette Champagnat'
    affiliations:
      - name: 'University of Washington'
        address: '1410 NE Campus Pkwy'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98195'
  - name: 'Abby Schamp'
    affiliations:
      - name: 'University of Washington'
        address: '1410 NE Campus Pkwy'
        city: 'Seattle, WA'
        state: 'WA'
        postal-code: '98195'
  - name: 'Samantha Schiano'
    affiliations:
      - name: 'ECS Federal in support of NMFS OST'
        address: 'East-West Hwy'
        city: 'Silver Spring, MD'
        state: 'MD'
        postal-code: '22031'
  - name: 'Fabio Prior Caltabellotta'
    affiliations:
      - name: 'Washington Department of Fish and Wildlife '
        address: '1111 Washington St SE'
        city: 'Olympia, WA'
        state: 'WA'
        postal-code: '98504-3150'
  - name: 'Vladlena Gertseva'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle'
        state: 'WA'
        postal-code: '98112-2097'
  - name: 'Ian Taylor'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle'
        state: 'WA'
        postal-code: '98112-2097'
  - name: 'Kiva Oken'
    affiliations:
      - name: 'NOAA Fisheries Northwest Fisheries Science Center'
        address: '2725 Montlake Boulevard East'
        city: 'Seattle'
        state: 'WA'
        postal-code: '98112-2097'
date: today
lang: en 
keep-tex: true 
mainfont: 'Latin Modern Sans' 
cover: support_files/Yelloweye_Rockfish.png
pdf-engine: lualatex
format: 
  pdf:
    knitr:
      opts_chunk:
        out.width: "\\textwidth" 
    code-block:
      out-width: "\\textwidth"
    documentclass: scrartcl
    number-sections: true
    template-partials:
      - 'support_files/before-body.tex'
      - 'support_files/_titlepage.tex'
    include-in-header:
      - 'support_files/in-header.tex'
      - 'support_files/pfmc.tex'
    include-after-body:
      text: |
        \newpage{}
        \printnoidxglossaries
    toc: true
    sansfont: 'Latin Modern Sans'
    lof: false
    lot: false
    titlepage-geometry: 
      - top=2in
      - bottom=1in
      - right=1in
      - left=1in
output-file: 'Yelloweye_rockfish_SAR_2025' 
params: 
   office: 'NWFSC'
   species: 'Yelloweye rockfish'
   spp_latin: 'Sebastes ruberrimus'
   region: 'U.S. West Coast'
bibliography: 
-  asar_references.bib 
---
\pagenumbering{roman}
\setcounter{page}{1}

\renewcommand{\thetable}{\roman{table}}
\renewcommand{\thefigure}{\roman{figure}}

```{r}
#| label: 'read_base_model'
#| eval: true
#| echo: false
#| message: false
#| warning: false

library(here)
library(dplyr)

source(here::here('Rcode/table_sens.R'))
source(here::here('Rcode/decision_table.R'))

base_mod <- 'model/2025_base_model'
mod_in <- r4ss::SS_read(here(base_mod))
mod_out <- r4ss::SS_output(here(base_mod), verbose = FALSE, printstats = FALSE)

```

```{r} 
#| label: 'output_and_quantities'
#| echo: false 
#| warning: false 
#| eval: false 
# load converted output from asar::convert_output() 
output <- utils::read.csv('/') 
# Call reference points and quantities below 
output <- output |> 
  dplyr::mutate(estimate = as.numeric(estimate), 
    uncertainty = as.numeric(uncertainty)) 
start_year <- as.numeric(min(output$year, na.rm = TRUE)) 
end_year <- (output |> 
  dplyr::filter(!(year %in% c('Virg', 'Init', 'S/Rcurve', 'INIT')), 
    !is.na(year)) |> 
  dplyr::mutate(year = as.numeric(year)) |> 
  dplyr::summarize(max_val = max(year)) |> 
  dplyr::pull(max_val))-10
# subset output to remove quantities that are split by factor 
output2 <- output |> 
  dplyr::filter(is.na(season), 
    is.na(fleet), 
    is.na(sex), 
    is.na(area), 
    is.na(growth_pattern), 
    is.na(subseason), 
    is.na(age))
# terminal fishing mortality 
Fend <- output2 |> 
  dplyr::filter(c(label == 'fishing_mortality' & year == end_year) | c(label == 'terminal_fishing_mortality' & is.na(year))) |>
  dplyr::pull(estimate)
# fishing mortality at msy 
# please change target if desired 
Ftarg <- output2 |>
  dplyr::filter(grepl('f_target', label) | grepl('f_msy', label) | c(grepl('fishing_mortality_msy', label) & is.na(year))) |>
  dplyr::pull(estimate)
# Terminal year F respective to F target 
F_Ftarg <- Fend / Ftarg
# terminal year biomass 
Bend <- output2 |>
  dplyr::filter(grepl('mature_biomass', label) | grepl('^biomass$', label),
    year == end_year) |>
  dplyr::pull(estimate)
# target biomass (msy) 
# please change target if desired 
Btarg <- output2 |>
  dplyr::filter(c(grepl('biomass', label) & grepl('target', label) & estimate >1) | label == 'biomass_msy') |>
  dplyr::pull(estimate)
# total catch in the last year 
total_catch <- output |>
  dplyr::filter(grepl('^catch$', label), 
  year == end_year,
    is.na(fleet),
    is.na(age),
    is.na(area),
    is.na(growth_pattern)) |>
  dplyr::pull(estimate)
# total landings in the last year 
total_landings <- output |>
  dplyr::filter(grepl('landings_weight', label), year == end_year,
    is.na(fleet),
    is.na(age)) |>
  dplyr::pull(estimate)
# spawning biomass in the last year
sbend <- output2 |>
  dplyr::filter(grepl('spawning_biomass', label), year == end_year) |>
  dplyr::pull(estimate) |>
    unique()
# overall natural mortality or at age 
M <- output |>
  dplyr::filter(grepl('natural_mortality', label)) |>
  dplyr::pull(estimate)
# Biomass at msy 
# to change to another reference point, replace msy in the following lines with other label 
Bmsy <- output2 |>
  dplyr::filter(c(grepl('biomass', label) & grepl('msy', label) & estimate >1) | label == 'biomass_msy') |>
  dplyr::pull(estimate)
# target spawning biomass(msy) 
# please change target if desired 
SBtarg <- output2 |>
  dplyr::filter(c(grepl('spawning_biomass', label) & grepl('msy$', label) & estimate >1) | label == 'spawning_biomass_msy$') |>
  dplyr::pull(estimate)
# steepness 
h <- output |> 
  dplyr::filter(grepl('steep', label)) |> 
  dplyr::pull(estimate)
# recruitment 
R0 <- output |> 
  dplyr::filter(grepl('R0', label) | grepl('recruitment_virgin', label)) |> 
  dplyr::pull(estimate)
# female SB (placeholder)

``` 

{{< pagebreak >}} 

Please cite this publication as: 

Johnston\*, M. A., Rosemond\*, R. C., Perl, E., Whitman, A., Barros, M., Champagnat, J., Schamp, A., Schiano, S., Prior Caltabellotta, F., Gertseva, V., Taylor, I., and Oken, K. (2025) Status of Yelloweye rockfish off the U.S. West Coast in 2025. Pacific Fishery Management Council. \pageref*{LastPage}{} pp.

\*These authors contributed equally to this work.

{{< pagebreak >}} 

\pagenumbering{roman}
\setcounter{page}{1}

\renewcommand{\thetable}{\roman{table}}
\renewcommand{\thefigure}{\roman{figure}}

```{r, results='asis'}
#| label: 'disclaimer'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('00a_disclaimer.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 
 
```{r, results='asis'}
#| label: 'load_tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('002_load_tables.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}}

```{r, results='asis'}
#| label: 'Executive Summary'
#| eval: true
#| echo: false
#| warning: false
 a <- knitr::knit_child('01_executive_summary.qmd', quiet = TRUE)
 cat(a, sep = '\n')
```

 {{< pagebreak >}} 

\setlength{\parskip}{5mm plus1mm minus1mm}
\pagenumbering{arabic}
\setcounter{page}{1}
\setcounter{section}{0}
\renewcommand{\thefigure}{\arabic{figure}}
\renewcommand{\thetable}{\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}


```{r, results='asis'}
#| label: 'introduction'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('02_introduction.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'data'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('03_data.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'assessment-configuration'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04a_model_summary.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

```{r, results='asis'}
#| label: 'assessment-sensitivity'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04b_model_results.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

```{r, results='asis'}
#| label: 'assessment-benchmarks'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('04c_model_diagnostics.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'discussion'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('05_management.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'acknowledgments'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('06_acknowledgments.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'references'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('07_references.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'tables'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('08_tables.qmd', quiet = TRUE)
cat(a, sep = '\n')
```

 {{< pagebreak >}} 

```{r, results='asis'}
#| label: 'figures'
#| eval: true
#| echo: false
#| warning: false
a <- knitr::knit_child('09_figures.qmd', quiet = TRUE)
cat(a, sep = '\n')
```
