---
title: "nwfscSurvey Vignette"
output:
  pdf_document: default
  html_document: default
date: "2025-01-23"
---
Set up: The first set of code is for the Bottom trawl survey
        The second set of code is for the triennial survey

Questions: is this a two area model as in WA/OR + CA? or WA + OR only? Is California a whole different assessment model?
If we are seeing catches in CA, what is the strata doing again?

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(nwfscSurvey)
dir.create(file.path(getwd(),"plots"))
```

## NWFSC.Combo Data

Initial pull and plotting of data: WCGBTS
```{r, warning = FALSE, message = FALSE}
catch = pull_catch(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo")

bio = pull_bio(
  common_name = "yelloweye rockfish", 
  survey = "NWFSC.Combo") |>
  dplyr::filter(Latitude_dd >= 42)

plot_cpue(
  catch = catch)

plot_bio_patterns(
  bio = bio, 
  col_name = "Length_cm")

wh_plot_proportion(
  data_catch = catch,
  data_bio = bio
)
```

### Calculate designed based index of abundance (not used in this assessment)

```{r, warning = FALSE, message = FALSE, eval = FALSE}
WCGBTS_strata <- CreateStrataDF.fn(
  names = c("shallow_OR", "deep_OR", "shallow_WA", "deep_WA"), 
  depths.shallow = c(55, 183, 55, 183),
  depths.deep    = c(183, 350, 183, 350),
  lats.south     = c(42, 42, 46, 46),
  lats.north     = c(46, 46, 49, 49)
)
```

```{r, warning = FALSE, message = FALSE, eval = FALSE}
biomass = get_design_based(
  data = catch,  
  strata = WCGBTS_strata,
  dir = file.path(dirname(dirname(getwd())), "Data", "NWFSC.Combo_and_Tri_design_based_indices"),
  printfolder = "")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE, eval = FALSE}
file.rename(from = file.path(dirname(dirname(getwd())), "Data", "NWFSC.Combo_and_Tri_design_based_indices","design_based_indices.csv"), 
            to = file.path(dirname(dirname(getwd())), "Data", "NWFSC.Combo_and_Tri_design_based_indices", "NWFSC.Combo_design_based_indices.csv"))
```

Plot coastwide
```{r, warning = FALSE, message = FALSE, eval = FALSE}
plot_index(
  data = biomass,
  plot = 1)
# thats only about a maximum of 200,500 fish.... very few...
```

Plot index for each strata
```{r, warning = FALSE, message = FALSE, eval = FALSE}
plot_index(
  data = biomass,
  plot = 2)
```

### Length composition data
```{r, warning = FALSE, message = FALSE}
# Expanded length comps are not used for the yelloweye assessment
# length_comps <- get_expanded_comps(
#     bio_data = bio,
#     catch_data = catch,
#     comp_bins = seq(10, 74, 2),
#     strata = WCGBTS_strata,
#     comp_column_name = "length_cm",
#     output = "full_expansion_ss3_format",
#     two_sex_comps = FALSE, #single sex model
#     input_n_method = "stewart_hamel")
# 
# plot_comps(
#   data = length_comps)

# Use raw length comps, you can tell this is what was used in the previous assessment because whole numbers are used
# The steward-hamel method is used to get the Nsamps, this is better than the simplified stewart-hamel approach used in the 2017 assessment and is fine to use in this years update according to the instructors.
# Nsamp method is the Stewart Hammel method where 
# Nsamp = n_tows + 0.0707 * n_fish when n_fish/n_tows < 55 and
# Nsamp = 4.89 * n_tows when n_fish/n_tows >= 55

raw_length_comps_n_trips <- get_raw_comps(
    data = bio,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("tows"),
    two_sex_comps = FALSE,
    fleet = 11,
    month = 7, 
    dir = getwd(),
    printfolder = "") 

raw_length_comps_n_trips <- raw_length_comps_n_trips$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_length_comps_n_fish <- get_raw_comps(
    data = bio,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("total_samples"),
    two_sex_comps = FALSE,
    fleet = 11,
    month = 7, 
    dir = getwd(),
    printfolder = "")

raw_length_comps_n_fish <- raw_length_comps_n_fish$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_length_comps <- raw_length_comps_n_trips |>
  left_join(raw_length_comps_n_fish, by = "year") |>
  dplyr::mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707*n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) |>
  write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "length_cm_unsexed_raw_10_74_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv"), row.names = FALSE)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
length_files <- list.files(path = getwd(), pattern = "length_cm")

file.rename(from = length_files[1],
            to = paste0("NWFSC.Combo_", length_files[1]))
file.rename(from = file.path(getwd(), paste0("NWFSC.Combo_", length_files[1])),
            to = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", paste0("NWFSC.Combo_", length_files[1])))

file.rename(from = length_files[2],
            to = paste0("NWFSC.Combo_", length_files[2]))
file.rename(from = file.path(getwd(), paste0("NWFSC.Combo_", length_files[2])),
            to = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", paste0("NWFSC.Combo_", length_files[2])))
```

### Marginal age composition data
```{r, warning = FALSE, message = FALSE}
# Expanded age comps are not used in the yelloweye assessment
# age_comps <- get_expanded_comps(
#     bio_data = bio,
#     catch_data = catch,
#     comp_bins = 0:65,
#     strata = WCGBTS_strata,
#     comp_column_name = "age",
#     output = "full_expansion_ss3_format",
#     two_sex_comps = FALSE, #single sex model
#     input_n_method = "stewart_hamel")
# 
# plot_comps(
#   data = age_comps)

raw_age_comps_n_trips <- get_raw_comps(
    data = bio,
    comp_bins = 0:65,
    input_n_method = "tows",
    comp_column_name = "age",
    fleet = 11,
    ageerr = 2,
    month = 7,
    dir = getwd(),
    printfolder = "", 
    two_sex_comps = FALSE) 

raw_age_comps_n_trips <- raw_age_comps_n_trips$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_age_comps_n_fish <- get_raw_comps(
    data = bio,
    comp_bins = 0:65,
    input_n_method = "total_samples",
    comp_column_name = "age",
    fleet = 11,
    ageerr = 2,
    month = 7,
    dir = getwd(),
    printfolder = "", 
    two_sex_comps = FALSE)

raw_age_comps_n_fish <- raw_age_comps_n_fish$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_age_comps <- raw_age_comps_n_trips |>
  left_join(raw_age_comps_n_fish, by = "year") |>
  mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707 * n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) |>
  write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "age_unsexed_raw_0_65_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv"), row.names = FALSE)

# plot_comps(
#   data = raw_age_comps_n_trips)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
age_files <- list.files(path = getwd(), pattern = "age_")

file.rename(from = age_files[1],
            to = paste0("NWFSC.Combo_", age_files[1]))
file.rename(from = file.path(getwd(), paste0("NWFSC.Combo_", age_files[1])), 
            to = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_age_comps", paste0("NWFSC.Combo_", age_files[1])))

file.rename(from = age_files[2],
            to = paste0("NWFSC.Combo_", age_files[2]))
file.rename(from = file.path(getwd(), paste0("NWFSC.Combo_", age_files[2])), 
          to = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_age_comps", paste0("NWFSC.Combo_", age_files[2])))
```

### Conditional age-at-length data
```{r, warning = FALSE, message = FALSE}
# The previous function used was actually deprecated and this function should have been used.
caal <- get_raw_caal(
  dat = bio, 
  len_bins = seq(10, 74, 2), 
  age_bins = 0:65,
  dir = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_CAAL"),
  printfolder = "")

# Reformat data to only have one sex
one_sex_caal <- caal |>
  dplyr::arrange(year, Lbin_lo) |>
  tidyr::pivot_longer(cols = f0:m65, names_to = "age", values_to = "count") |>
  dplyr::mutate(input_n = case_when(
    grepl("m", age) ~ 0,
    grepl("f", age) ~ input_n
  )) |>
  mutate(age = as.numeric(stringr::str_remove_all(age, "f|m"))) |>
  dplyr::group_by(year, Lbin_lo, age) |>
  dplyr::summarize(month = 7,
                   fleet = 11,
                   sex = 0,
                   partition = unique(partition),
                   ageerr = 2,
                   Lbin_hi = unique(Lbin_hi),
                   Nsamp = sum(input_n),
                   count = sum(count)) |>
  dplyr::select(year, month, fleet, sex, partition, ageerr, Lbin_lo, Lbin_hi, Nsamp, age, count) |>
  tidyr::pivot_wider(names_from = "age", values_from = "count")

write.csv(one_sex_caal, file = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_CAAL", "processed_one_sex_caal.csv"), row.names = FALSE)

# PlotMap.fn(
#   dat = catch)
```

### Weight-Length Relationship
```{r, warning = FALSE, message = FALSE}
# Ian said we can just use the nwfsc.combo data and don't need to filter by the same latitudes as the index was, just use all data
wt_len_est <- estimate_weight_length(
  bio,
  col_length = "Length_cm",
  col_weight = "Weight_kg",
  verbose = FALSE
)

wt_len_est

plot_weight_length(
  bio,
  dir = file.path(dirname(getwd()), "model_params"),
  estimates = wt_len_est,
  col_length = "length_cm",
  col_weight = "weight_kg",
  two_sex = FALSE,
  add_save_name = NULL,
  height = 7,
  width = 7,
  dpi = 300
)
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}
if(dir.exists(file.path(getwd(),"NWFSC.Combo_plots"))){
  plots <- list.files(file.path(getwd(), "plots"))
  lapply(plots, function(x){
    file.rename(from = file.path(getwd(), "plots", plots[x]),
                to = file.path(getwd(), "NWFSC.Combo_plots", plots[x]))
  })
} else {
  file.rename(from = file.path(getwd(), "plots"), 
            to = file.path(getwd(), "NWFSC.Combo_plots"))
}
```

## Triennial Data
```{r, echo = FALSE, message = FALSE, warning = FALSE}
if(dir.exists(file.path(getwd(),"plots")) == FALSE) {
  dir.create(file.path(getwd(),"plots"))
}
```

Initial pull and plotting of data: Triennial
```{r, warning = FALSE, message = FALSE}
catchT = pull_catch(
  common_name = "yelloweye rockfish", 
  survey = "Triennial")

#this is creating a list, just use $age_data
bioT = pull_bio(
  common_name = "yelloweye rockfish", 
  survey = "Triennial")

bioT <- bioT$length_data |>
  dplyr::filter(Latitude_dd >= 42) # Ian said to use length data

plot_cpue(
  catch = catchT)

plot_bio_patterns(
  bio = bioT, 
  col_name = "Length_cm")

wh_plot_proportion(
  data_catch = catchT,
  data_bio = bioT
)
```

### Calculate designed based index of abundance

```{r, warning = FALSE, message = FALSE}
Triennial_strata = CreateStrataDF.fn(names=c("OR", "WA"), 
                           depths.shallow = c(55, 55),
                           depths.deep    = c(350, 350),
                           lats.south     = c(42, 46),
                           lats.north     = c(46, 49))
```

```{r, warning = FALSE, message = FALSE}
biomassT = get_design_based(
  data = catchT,  
  strata = Triennial_strata,
  dir = file.path(dirname(dirname(getwd())), "Data", "NWFSC.Combo_and_Tri_design_based_indices"),
  printfolder = "")
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
file.rename(from = file.path(dirname(dirname(getwd())), "Data", "NWFSC.Combo_and_Tri_design_based_indices","design_based_indices.csv"), 
            to = file.path(dirname(dirname(getwd())), "Data", "NWFSC.Combo_and_Tri_design_based_indices", "Tri_design_based_indices.csv"))
```

Plot coastwide
```{r, warning = FALSE, message = FALSE}
plot_index(
  data = biomassT,
  plot = 1)
# not a lot of fish...
```

Plot index for each strata
```{r, warning = FALSE, message = FALSE}
plot_index(
  data = biomassT,
  plot = 2)
```

### Length composition data
```{r, warning = FALSE, message = FALSE}
# Expanded length comps are not used for the yelloweye assessment
# length_compsT <- get_expanded_comps(
#     bio_data = bioT,
#     catch_data = catchT,
#     comp_bins = seq(10, 74, 2),
#     strata = Triennial_strata,
#     comp_column_name = "length_cm",
#     output = "full_expansion_ss3_format",
#     two_sex_comps = FALSE, #single sex model
#     input_n_method = "stewart_hamel")
# 
# plot_comps(
#   data = length_compsT)

raw_length_comps_n_trips <- get_raw_comps(
    data = bioT,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("tows"),
    two_sex_comps = FALSE,
    fleet = 10,
    month = 7,
    dir = getwd(),
    printfolder = "")

raw_length_comps_n_trips <- raw_length_comps_n_trips$unsexed |>
  dplyr::rename(n_tows = input_n)

raw_length_comps_n_fish <- get_raw_comps(
    data = bioT,
    comp_bins = seq(10, 74, 2),
    comp_column_name = "length_cm",
    input_n_method = c("total_samples"),
    two_sex_comps = FALSE,
    fleet = 10,
    month = 7, 
    dir = getwd(),
    printfolder = "")

raw_length_comps_n_fish <- raw_length_comps_n_fish$unsexed |>
  dplyr::rename(n_samps = input_n)|>
  dplyr::select(year, n_samps)

raw_length_comps <- raw_length_comps_n_trips |>
  left_join(raw_length_comps_n_fish, by = "year") |>
  dplyr::mutate(
    n_samps_per_tow = n_samps / n_tows,
    Nsamp = round(ifelse(n_samps_per_tow < 55, (n_tows + 0.0707*n_samps), 4.89*n_tows), digits = 2)) |>
  dplyr::select(year, month, fleet, sex, partition, Nsamp, everything(), -n_samps_per_tow, -n_samps, -n_tows) |>
  write.csv(file.path(dirname(dirname(getwd())), "Rcode", "survey_length_age", "length_cm_unsexed_raw_10_74_yelloweye rockfish_groundfish_triennial_shelf_survey.csv"), row.names = FALSE)

# plot_comps(
#   data = raw_length_comps_n_trips)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
length_files <- list.files(path = getwd(), pattern = "length_cm")

file.rename(from = length_files[1],
            to = paste0("Tri_", length_files[1]))
file.rename(from = file.path(getwd(), paste0("Tri_", length_files[1])),
            to = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", paste0("Tri_", length_files[1])))

file.rename(from = length_files[2],
            to = paste0("Tri_", length_files[2]))
file.rename(from = file.path(getwd(), paste0("Tri_", length_files[2])),
            to = file.path(dirname(dirname(getwd())), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", paste0("Tri_", length_files[2])))
```

### No marginal age composition data for triennial survey
```{r, warning = FALSE, message = FALSE}
#no age data from Triennial survey?

#age_compsT <- get_expanded_comps(
#    bio_data = bioT,
#    catch_data = catchT,
#    comp_bins = 0:65,
#    strata = Triennial_strata,
#    comp_column_name = "age",
#    output = "full_expansion_ss3_format",
#    two_sex_comps = FALSE, #single sex model
#    input_n_method = "stewart_hamel")

#plot_comps(
#  data = age_compsT)

#raw_age_compsT <- get_raw_comps(
#    data = bioT,
#    comp_bins = 0:65,
#    comp_column_name = "age",
#    two_sex_comps = FALSE) #single sex model
```

### No conditional age-at-length data for triennial survey
```{r, warning = FALSE, message = FALSE}
#no age data from Triennial survey?

#caalT <- SurveyAgeAtLen.fn(
#  datAL = bioT, 
#  datTows = catchT,
#  strat.df = Triennial_strata,
#  lgthBins = seq(10, 74, 2), 
#  ageBins = 0:65)

PlotMap.fn(
  dat = catchT)
```

```{r, echo = FALSE, message = FALSE, warning = FALSE}
if(dir.exists(file.path(getwd(),"Tri_plots"))){
  plots <- list.files(file.path(getwd(), "plots"))
  lapply(plots, function(x){
    file.rename(from = file.path(getwd(), "plots", plots[x]),
                to = file.path(getwd(), "Tri_plots", plots[x]))
  })
} else {
  file.rename(from = file.path(getwd(), "plots"), 
            to = file.path(getwd(), "Tri_plots"))
}
```

## WCGOP Fisheries Discards
```{r, warning = FALSE, message = FALSE, }
gemm_data <- pull_gemm("yelloweye rockfish", 
                       dir = file.path(dirname(dirname(getwd())), "Data", "raw"), 
                       verbose = TRUE)

head(gemm_data, n = 20)
```


## Composition comparison plots
### NWFSC.Combo Comparisons
```{r}
# Lengths
nwfsc_lengths_new <- read.csv(file.path(getwd(), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", "NWFSC.Combo_length_cm_unsexed_raw_10_74_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv")) |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, input_n, `u10`:`u74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -input_n), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("u", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    length = as.numeric(length)
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path(getwd(), "model", "2017_yelloweye_model_updated_ss3_exe"))

nwfsc_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 11)|>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    length = as.numeric(length)
  )

together <- rbind(nwfsc_lengths_old, nwfsc_lengths_new)
comparison_plot <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5))
ggsave(plot = comparison_plot, "nwfsc_length_comp_comparisons.png", path = file.path(getwd(), "Rcode", "survey_length_age", "NWFSC.Combo_plots"))

# Ages
nwfsc_ages_new <- read.csv(file.path(getwd(), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", "NWFSC.Combo_age_unsexed_raw_0_65_yelloweye rockfish_groundfish_slope_and_shelf_combination_survey.csv")) |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, input_n, `u0`:`u65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -input_n), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("u", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    age = as.numeric(age)
  )

nwfsc_ages_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == 11)|>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    age = as.numeric(age)
  )

together <- rbind(nwfsc_ages_old, nwfsc_ages_new)
comparison_plot <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5))
ggsave(plot = comparison_plot, "nwfsc_age_comp_comparisons.png", path = file.path(getwd(), "Rcode", "survey_length_age", "NWFSC.Combo_plots"))
```

### Triennial length comparisons only
```{r}
# Lengths
tri_lengths_new <- read.csv(file.path(getwd(), "Data", "processed", "NWFSC.Combo_and_Tri_length_comps", "Tri_length_cm_unsexed_raw_10_74_yelloweye rockfish_groundfish_triennial_shelf_survey.csv")) |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, input_n, `u10`:`u74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -input_n), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("u", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    length = as.numeric(length)
  )


tri_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 10)|>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    # NsampsFreq = freq/unique(Nsamps), still waiting on these
    length = as.numeric(length)
  )

together <- rbind(tri_lengths_old, tri_lengths_new)
comparison_plot <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5))
ggsave(plot = comparison_plot, "tri_length_comp_comparisons.png", path = file.path(getwd(), "Rcode", "survey_length_age", "Tri_plots"))
```


### Age composition ccomparisons (NWFSC.Combo only)
Both of these use csv files that were formatted to input in the SAC tool with different numbered fleets to represent the years but 
you could easily extract data from the previous assessment and rbind it with the updated data. The big thing you would want is 
```{r}
nwfsc_ages<-read.csv(file.path(getwd(), "NWFSC.Combo_plots", "nwfsc.combo_comparison_ages.csv"))

long_nwfsc_ages <- nwfsc_ages |>
              dplyr::mutate(type = dplyr::case_when(Fleet == 1 ~ "previous assessment",
                                      Fleet == -1 ~ "previous assessment",
                                      Fleet == 2 ~ "current assessment",
                                      Fleet == -2 ~ "current assessment"),
                     age_type = dplyr::case_when(Fleet == 1 ~ "caal",
                                      Fleet == -1 ~ "maal",
                                      Fleet == 2 ~ "caal",
                                      Fleet == -2 ~ "maal"))|>
                    dplyr::select(Year, type, age_type, `X0`:`X65`) |>
                    tidyr::pivot_longer(cols = c(-Year, -type, -age_type), names_to = 'age', values_to = 'freq') |>
                    dplyr::mutate(age = gsub("X","",age))
          
old_comps <- long_nwfsc_ages |>
             dplyr::filter(type == "previous assessment") |>
             dplyr::filter(age_type == "maal") |>
             dplyr::group_by(Year) |>
             dplyr::mutate(freq = freq/sum(freq),
                    age = as.numeric(age)) 
new_comps <- long_nwfsc_ages |>
             dplyr::filter(type == "current assessment") |>
             dplyr::filter(age_type == "maal")|>
             dplyr::group_by(Year) |>
             dplyr::mutate(freq = freq/sum(freq),
                    age = as.numeric(age)) 

# Kiva's method of looking at the age comps...it didn't make sense to my brain
# dplyr::bind_rows(list(new = new_comps,
#                old = old_comps),
#           .id = 'type') |> 
#   tidyr::pivot_wider(names_from = type, values_from = freq) |>
#   ggplot2::ggplot(aes(x = new, y = old, col = age)) +
#   ggplot2::geom_point()

together <- rbind(old_comps, new_comps)
comparison_plot <- together |>
                   dplyr::filter(freq > 0 ) |>
                   ggplot2::ggplot(aes(x = Year, y = age, col = type, size = freq)) +
                   ggplot2::geom_point(position=position_dodge(0.5))
ggsave(plot = comparison_plot, "nwfsc_age_comp_comparisons.png", path = file.path(getwd(), "NWFSC.Combo_plots"))

```

### Length composition comparisons 
```{r}
nwfsc_lengths<-read.csv(file.path(getwd(), "NWFSC.Combo_plots", "nwfsc.combo_comparison_lengths.csv"))

long_nwfsc_lengths <- nwfsc_lengths |>
              dplyr::mutate(type = dplyr::case_when(Fleet == 1 ~ "previous assessment",
                                      Fleet == 2 ~ "current assessment")) |>
              dplyr::select(Year, type, Nsamps, `X10`:`X74`) |>
              tidyr::pivot_longer(cols = c(-Year, -type, -Nsamps), names_to = 'length', values_to = 'freq') |>
              dplyr::mutate(length = gsub("X","",length))

old_comps <- long_nwfsc_lengths |>
             dplyr::filter(type == "previous assessment") |>
             dplyr::group_by(Year) |>
             dplyr::mutate(freq = freq/sum(freq),
                    NsampsFreq = freq/unique(Nsamps),
                    length = as.numeric(length)) 
new_comps <- long_nwfsc_lengths |>
             dplyr::filter(type == "current assessment") |>
             dplyr::group_by(Year) |>
             dplyr::mutate(freq = freq/sum(freq),
                    NsampsFreq = freq/unique(Nsamps),
                    length = as.numeric(length)) 

together <- rbind(old_comps, new_comps)
comparison_plot <- together |>
                   dplyr::filter(freq > 0 ) |>
                   ggplot2::ggplot(aes(x = Year, y = length, col = type, size = freq)) +
                   ggplot2::geom_point(position=position_dodge(0.5))
ggsave(plot = comparison_plot, "nwfsc_length_comp_comparisons.png", path = file.path(getwd(), "NWFSC.Combo_plots"))

## Triennial length compositions
tri_lengths <- read.csv(file.path(getwd(), "Tri_plots", "triennial_comparison_lengths.csv"))
long_tri_lengths <- tri_lengths |>
              dplyr::mutate(type = dplyr::case_when(Fleet == 1 ~ "previous assessment",
                                      Fleet == 2 ~ "current assessment")) |>
              dplyr::select(Year, type, Nsamps, `X10`:`X74`) |>
              tidyr::pivot_longer(cols = c(-Year, -type, -Nsamps), names_to = 'length', values_to = 'freq') |>
              dplyr::mutate(length = gsub("X","",length))

old_comps <- long_tri_lengths |>
             dplyr::filter(type == "previous assessment") |>
             dplyr::group_by(Year) |>
             dplyr::mutate(freq = freq/sum(freq),
                    NsampsFreq = freq/unique(Nsamps),
                    length = as.numeric(length)) 
new_comps <- long_tri_lengths |>
             dplyr::filter(type == "current assessment") |>
             dplyr::group_by(Year) |>
             dplyr::mutate(freq = freq/sum(freq),
                    NsampsFreq = freq/unique(Nsamps),
                    length = as.numeric(length)) 

together <- rbind(old_comps, new_comps)
comparison_plot <- together |>
                   dplyr::filter(freq > 0 ) |>
                   ggplot2::ggplot(aes(x = Year, y = length, col = type, size = freq)) +
                   ggplot2::geom_point(position=position_dodge(0.5))
ggsave(plot = comparison_plot, "tri_length_comp_comparisons.png", path = file.path(getwd(), "Tri_plots"))
```