---
title: "Yelloweye Commercial Catch"
author: Samantha Schiano
format: html
---

## Previous Assessment Calcs

<!--- - CA historial (1916-1980) commercial catches were recontructed
 - used spatial info on aggregate rockfish catches back to 1916 and used intermittent species composition records by markey category to apportion the catches in mixed species market categories to species, fishing gears, and ports
 - Linear rampoing was applied over the period btwn 1889-1916
 --->

-   landings from 1981-2016 were obtained from PacFIN (extracted by hear type and state), then combined into the area-specific fishing fleets used in the assessment
-   Supplement the year-specific landings from PacFIN with discard amounts estimated by WCGOP = total commercial catch

### Previous data sources needed

-   PacFIN catches
-   WCGOP

CA - from PacFIN OR - reconstructed landings through 1986; recent landings (1987-2017) obtained from PacFIN OR pacfin landings from 1987-1999 were supplemented with add'l estimates from landing\`s reported w/in rockfish market categories

Area-specific fishing fleets: California Oregon-Washington

## Relevant Notes from GitHub Discussions

-   For yelloweye, the bottom and midwater trawls are in the same fleet, don't worry about that.

Often the gear groupings for TWL and TWS (shrimp trawl) are combined, and then everything else goes into non-trawl. You might check with people analyzing the catch data to make sure that is the case. I have seen shrimp trawl catches put into non-trawl fleets before. Whatever you do should be consistent between the composition data and the landings.

-   cleanPacFIN() is taking care of these HKL and TWL cases

-   PacFIN has a lot of different users and three different state agencies that provide data. As a result, there are a LOT of different columns to denote gear. For stock assessment, we most commonly use PACFIN_GROUP_GEAR_CODE. This is only available for the catch data-- but you can see which specific PACFIN_GEAR_CODE matches up with the more general PACFIN_GROUP_GEAR_CODE.

-   use the same data provided in the last assessment then all the years after will be sources from PacFIN -- follow up with Fabio (WA) and Ali (OR) to talk about commercial catches (do we also need to ask julia? (CA)) -- Is the cv assumed or calculated? Looks like the previous assessment cv's are just set to 0.01

## Load in data

```{r}
#| echo: false
common_name <- "yelloweye"
species_code <- "YEYE"

or <- read.csv(file.path(getwd(), "data", "Oregon Commercial landings_457_2024.csv")) |>
  dplyr::mutate(STATE = "OR",
                FLEET = dplyr::case_when(
                  FLEET == "NTRW" ~ "NONTWL",
                  FLEET == "TRW" ~ "TWL",
                  TRUE ~ NA
                  ))
                
file_catch <- fs::path(getwd(), "data", "confidential", "PacFIN.YEYE.CompFT.12.Dec.2024.RData")
# file_bds <- fs::path(getwd(), "data", "confidential", "PacFIN.YEYE.bds.12.Dec.2024.RData")
# head(catch.pacfin)
```


Plotting raw data


Load in data now

```{r}
#| echo: false
load(file_catch) # catch.pacfin
```

## Catch Data

1.  Plot landings by state, gear, and fleet
2.  Compare current landings with those in the last assessment
3.  Make sample size tables to determine confidentiality needs


Identify gears in pacfin to figure out trawl and non-trawl fleets
```{r}
#| echo: false
#| eval: true
#| output: false
colnames(catch.pacfin)

catch.pacfin |> 
  dplyr::select(PACFIN_GROUP_GEAR_CODE,PACFIN_GEAR_CODE) |> 
  dplyr::distinct() |>
  dplyr::arrange(PACFIN_GROUP_GEAR_CODE)

unique(catch.pacfin$PACFIN_SPECIES_CODE)

pacfin_gears <- pacfintools::get_codelist() |> dplyr::select(GRID, GROUP)
```

### Plots

```{r}
#| echo: false
#| warning: false
#| message: false
library(ggplot2)

# summarize pacfin data by gear code
catch_gear <- catch.pacfin |>
  dplyr::left_join(pacfin_gears, by = c("PACFIN_GEAR_CODE" = "GRID")) |>
  dplyr::mutate(FLEET = dplyr::case_when(
    GROUP == "TWL" ~ "TWL",
    GROUP == "TWS" ~ "TWL",
    TRUE ~ "NONTWL"
  )) |>
  dplyr::group_by(LANDING_YEAR, FLEET, PACFIN_GEAR_CODE, COUNTY_STATE) |> 
  dplyr::summarise(TOTAL_CATCH = sum(LANDED_WEIGHT_MTONS)) |>
  dplyr::ungroup() |>
  dplyr::filter(!is.na(TOTAL_CATCH))

ggplot(data = catch_gear, aes(x = LANDING_YEAR, y = TOTAL_CATCH, fill = PACFIN_GEAR_CODE)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ COUNTY_STATE)
```

```{r}
#| echo: false
#| warning: false
#| message: false
# summarize pacfin data by state and assign gear codes to trawl
catch_state <- catch.pacfin |>
  dplyr::left_join(pacfin_gears, by = c("PACFIN_GEAR_CODE" = "GRID")) |>
  dplyr::mutate(FLEET = dplyr::case_when(
    GROUP == "TWL" ~ "TWL",
    GROUP == "TWS" ~ "TWL",
    TRUE ~ "NONTWL"
  )) |>
  dplyr::group_by(LANDING_YEAR, FLEET, COUNTY_STATE) |> 
  dplyr::summarise(TOTAL_CATCH = sum(LANDED_WEIGHT_MTONS)) |>
  dplyr::ungroup() |>
  na.omit()

ggplot(data = catch_state, aes(x = LANDING_YEAR, y = TOTAL_CATCH, fill = FLEET)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ COUNTY_STATE) +
  labs(
    x="Year",
    y="Catch (mt)",
    fill="Fleet"
  ) +
  theme_classic()

ggplot(data = catch_state, aes(x = LANDING_YEAR, y = TOTAL_CATCH, fill = COUNTY_STATE)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ FLEET) +
  labs(
    x="Year",
    y="Catch (mt)",
    fill="State"
  ) +
  theme_classic()

```

### Plots for recent years

Data is same as above, just subset for update years

```{r}
#| warning: false
#| message: false
ggplot(data = subset(catch_state, LANDING_YEAR>2015), aes(x = as.factor(LANDING_YEAR), y = TOTAL_CATCH, fill = FLEET)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ COUNTY_STATE) +
  labs(
    x="Year",
    y="Catch (mt)",
    fill="Fleet"
  ) +
  theme_classic()

ggplot(data = subset(catch_state, LANDING_YEAR>2015), aes(x = as.factor(LANDING_YEAR), y = TOTAL_CATCH, fill = COUNTY_STATE)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  facet_wrap(~ FLEET) +
  labs(
    x="Year",
    y="Catch (mt)",
    fill="State"
  ) +
  theme_classic()
```

## Summarize Catch data by state, year, and gear

This is the summarized data for use in the update assessment

```{r}
#| echo: false
#| warning: false
#| message: false
catch <- catch.pacfin |>
  dplyr::left_join(pacfin_gears, by = c("PACFIN_GEAR_CODE" = "GRID")) |>
  dplyr::mutate(FLEET = dplyr::case_when(
    GROUP == "TWL" ~ "TWL",
    GROUP == "TWS" ~ "TWL",
    TRUE ~ "NONTWL"
  )) |>
  dplyr::group_by(LANDING_YEAR, FLEET, COUNTY_STATE) |>
  dplyr::summarise(TOTAL_CATCH = sum(LANDED_WEIGHT_MTONS)) |>
  dplyr::rename(YEAR = LANDING_YEAR,
                TOTAL = TOTAL_CATCH,
                STATE = COUNTY_STATE
  ) |>
  dplyr::ungroup() |>
  dplyr::select(YEAR, FLEET, STATE, TOTAL) |>
  dplyr::filter(!is.na(STATE)
                # filter data to above 2016 since we are only adding this into it
                # YEAR > 2015
                )
```

## More Plotting

```{r}
#| echo: false
#| warning: false
#| message: false
catch_plt <- ggplot(data = catch) +
  geom_line(aes(x = YEAR, y = TOTAL, color = FLEET), linewidth = 1) +
  facet_wrap(~ STATE) +
  ggplot2::labs(
    x = "Year",
    y = "Catch (mt)",
    color = "Fleet"
  )
stockplotr::add_theme(catch_plt)
```

Look at WA Landings in 2023

```{r}
#| warning: false
#| message: false
WA.catch <- catch.pacfin |>
    dplyr::filter(LANDING_YEAR == 2023,
                  COUNTY_STATE == "WA")
max(WA.catch$LANDED_WEIGHT_MTONS)
chk <- WA.catch |>
  dplyr::select(LANDED_WEIGHT_MTONS, LANDING_YEAR, COUNTY_STATE, PACFIN_GEAR_CODE)
CHK2 <- WA.catch[c(57,128),]

WA_catch_gear <- catch.pacfin |>
  dplyr::left_join(pacfin_gears, by = c("PACFIN_GEAR_CODE" = "GRID")) |>
  dplyr::mutate(FLEET = dplyr::case_when(
    GROUP == "TWL" ~ "TWL",
    TRUE ~ "NONTWL"
  )) |>
  dplyr::group_by(LANDING_YEAR, FLEET, PACFIN_GEAR_CODE, COUNTY_STATE) |> 
  dplyr::summarise(TOTAL_CATCH = sum(LANDED_WEIGHT_MTONS)) |>
  dplyr::ungroup() |>
  dplyr::filter(!is.na(TOTAL_CATCH),
                COUNTY_STATE == "WA")

ggplot(data = WA_catch_gear, aes(x = LANDING_YEAR, y = TOTAL_CATCH, fill = PACFIN_GEAR_CODE)) +
  geom_bar(stat = "identity", position = "stack", color = "black") +
  labs(
    x="Year",
    y="Catch (mt)",
    fill="Gear"
  ) +
  ggtitle("Washington Yelloweye Commercial Catch")+
  theme_light()
```

### Comparison to OR state data

Pacfin OR from 2016-present vs OR state catch data from 2016-present

```{r}
#| eval: true
#| warning: false
#| message: false
# Does not work
# Cath data subset to >2015 (chose 2015 since the 2016 data is missing for CA nontwl)
pacfin.or <- catch |>
  dplyr::filter(STATE == "OR") |>
  dplyr::mutate(SOURCE = "PACFIN")
or_compar <- or |> dplyr::select(YEAR, FLEET, TOTAL, STATE) |>
  dplyr::mutate(SOURCE = "STATE") |>
  rbind(pacfin.or) |>
  dplyr::filter(YEAR > 2015)
  
ggplot(data = or_compar, aes(x = as.factor(YEAR), y = TOTAL, fill = SOURCE)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ FLEET) +
  labs(
    x="Year",
    y="Catch (mt)",
    fill="Source"
  )
```

Checking the difference of catches reported by PacFIN vs OR state in 2018 due to discrepancy above

```{r}
#| echo: false
#| eval: false
#| warning: false
#| message: false
# difference btwn pacfin and OR state commercial data in 2018
pf_num <- pacfin.or |>
  dplyr::filter(YEAR == 2018,
                FLEET == "NONTWL") |>
  dplyr::pull(TOTAL)
pull_or_2018 <- or_compar |>
  dplyr::filter(YEAR == 2018,
                FLEET == "NONTWL")
diff <- pull_or_2018$TOTAL[pull_or_2018$SOURCE=="STATE"] - pull_or_2018$TOTAL[pull_or_2018$SOURCE=="PACFIN"]
diff

chk_2018_nontwl <- or |>
  dplyr::filter(YEAR == 2018,
                FLEET == "NONTWL")
chk_pac_2018_nontwl <- catch.pacfin
```

Result of above leads us to use the OR state data to be more convervative even though it's minimal on the grand scheme of things. 


## Final Catch

```{r}
#| echo: false
#| warning: false
#| message: false
# due to above results - replace OR fleet data from PacFIN with OR state data
or2 <- or |>
  dplyr::select(YEAR, FLEET, TOTAL, STATE) |>
  dplyr::filter(YEAR > 2015)

# remove OR data from catch
pacfin.cawa <- catch |>
  dplyr::filter(STATE != "OR")

# combine subset PacFIN catch data (WA and CA) and OR state data
catch2 <- rbind(pacfin.cawa, or2)

# read in commercial discards
discards <- read.csv(file.path(getwd(), "commercial_discards.csv")) |> 
  dplyr::select(-X)

# Combine OR and WA discards for fleet
discards_ORWA <- discards |>
  dplyr::filter(grepl("OR|WA", fleet)) |>
  tidyr::separate_wider_delim(fleet, delim = "-", names = c("fleet", "state")) |>
  dplyr::group_by(year, fleet) |>
  dplyr::summarise(discards = sum(total_discards)) |>
  dplyr::ungroup() |>
  dplyr::mutate(state = "ORWA")

# Combine CA and ORWA discards into final discards df
discards_all <- discards |>
  dplyr::filter(grepl("CA", fleet)) |>
  tidyr::separate_wider_delim(fleet, delim = "-", names = c("fleet", "state")) |>
  dplyr::rename(discards = total_discards) |>
  rbind(discards_ORWA) |>
  dplyr::mutate(ST_FLEET = glue::glue("{state}_{fleet}")) |>
  dplyr::select(-c(fleet, state)) |>
  dplyr::filter(year > 2015)

# Combine WA and OR fleets
catch_orwa <- catch2 |>
  dplyr::filter(STATE %in% c("WA", "OR")) |>
  dplyr::group_by(YEAR, FLEET) |>
  dplyr::summarise(CATCH = sum(TOTAL)) |>
  dplyr::ungroup() |>
  dplyr::mutate(STATE = "ORWA")

# Subset catch to just CA and add back in ORWA then restrucutre DF for final output
comm_catch <- catch2 |>
  dplyr::filter(STATE == "CA") |>
  dplyr::rename(CATCH = TOTAL) |>
  rbind(catch_orwa) |>
  dplyr::mutate(ST_FLEET = glue::glue("{STATE}_{FLEET}")) |>
  dplyr::filter(YEAR > 2015) |>
  dplyr::left_join(discards_all, by = c("YEAR" = "year", "ST_FLEET" = "ST_FLEET")) |>
  dplyr::mutate(discards = dplyr::case_when(is.na(discards) ~ 0,
                                            TRUE ~ discards),
                CATCH = dplyr::case_when(is.na(CATCH) ~ 0,
                                            TRUE ~ CATCH),
                TOTAL_CATCH = CATCH + discards,
                SEAS = "1",
                CATCH_SE = 0.01) |>
  dplyr::select(YEAR, SEAS, ST_FLEET, TOTAL_CATCH, CATCH_SE)

# export catches
# write.csv(comm_catch, file = file.path(getwd(), "yelloweye_commercial_catch_2016_2024.csv"), append = FALSE)
```

Sample size tables

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
# 9,241 occruances of vessels having <=3 trips in a given year
vessel_yr_n <- catch.pacfin |>
  dplyr::group_by(LANDING_YEAR, VESSEL_ID) |>
  dplyr::summarize(total_count=dplyr::n())

vessel_yr_n |>
  dplyr::filter(total_count <= 3)

# 2,556 vessels with 3 or less trips over all years
vessel_n <- catch.pacfin |>
  dplyr::group_by(VESSEL_ID) |>
  dplyr::summarize(total_count=dplyr::n())
vessel_n |>
  dplyr::filter(total_count <= 3)

# no years with occurances less than 3
year_n <- catch.pacfin |>
  dplyr::group_by(LANDING_YEAR) |>
  dplyr::summarize(total_count=dplyr::n())
year_n |>
  dplyr::filter(total_count <= 3)
```

Vessel count from for summarizing data for catches (2017-present)

```{r}
#| echo: false
#| warning: false
#| message: false
catch <- catch.pacfin |>
  dplyr::left_join(pacfin_gears, by = c("PACFIN_GEAR_CODE"="GRID")) |>
  dplyr::mutate(FLEET = dplyr::case_when(
    GROUP == "TWL" ~ "TWL",
    TRUE ~ "NONTWL"
  )) |>
  dplyr::group_by(LANDING_YEAR, FLEET, COUNTY_STATE) |>
  dplyr::summarise(TOTAL_CATCH = sum(LANDED_WEIGHT_MTONS)) |>
  dplyr::rename(YEAR = LANDING_YEAR,
                TOTAL = TOTAL_CATCH,
                STATE = COUNTY_STATE
  ) |>
  dplyr::ungroup() |>
  dplyr::select(YEAR, FLEET, STATE, TOTAL) |>
  dplyr::filter(!is.na(STATE),
                # filter data to above 2016 since we are only adding this into it
                YEAR > 2015)

trips <- catch.pacfin |>
  dplyr::left_join(pacfin_gears, by = c("PACFIN_GEAR_CODE"="GRID")) |>
  dplyr::mutate(FLEET = dplyr::case_when(
    GROUP == "TWL" ~ "TWL",
    TRUE ~ "NONTWL"
  )) |>
  dplyr::filter(LANDING_YEAR > 2015) |>
  dplyr::group_by(LANDING_YEAR, FLEET, COUNTY_STATE) |>
  dplyr::summarise(count = dplyr::n()) |>
  dplyr::filter(!is.na(COUNTY_STATE))
  
```

Plot OR Data

```{r}
#| echo: false
#| eval: true 
#| warning: false
#| message: false
or_land <- ggplot2::ggplot(data = or2) +
  ggplot2::geom_line(ggplot2::aes(x = YEAR, y = TOTAL, color = FLEET))
satf::add_theme(or_land)
```

## Catches compared to the last assessment

```{r}
#| echo: false
#| warning: false
#| message: false
# from r4ss
rich.colors.short <- function(n, alpha = 1) {
  x <- seq(0, 1, length = n)
  r <- 1 / (1 + exp(20 - 35 * x))
  g <- pmin(pmax(0, -0.8 + 6 * x - 5 * x^2), 1)
  b <- dnorm(x, 0.25, 0.15) / max(dnorm(x, 0.25, 0.15))
  rgb.m <- matrix(c(r, g, b), ncol = 3)
  rich.vector <- apply(rgb.m, 1, function(v) rgb(v[1], v[2], v[3], alpha = alpha))
}
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
# load in input catches
hist_c <- read.csv(file.path(getwd(), "data", "ss3_input_comm_catch.csv")) |>
  dplyr::select(YEAR, SEAS, ST_FLEET, CATCH, CATCH_SE) |>
  dplyr::rename(TOTAL_CATCH = CATCH)

all_catch <- rbind(hist_c, comm_catch) |>
  dplyr::mutate(
    # YEAR = as.factor(YEAR),
    CATCH = as.numeric(TOTAL_CATCH)
    )

# identify colors for plot
all_fleets <- c("CA_TWL", "CA_NONTWL", "CA_REC", "ORWA_TWL", "ORWA_NONTWL", "OR_REC", "WA_REC")
colors <- rich.colors.short(length(unique(all_fleets)) + 1)[-1]

fleetcols <- c("CA_TWL" = "#0000CBFF", "CA_NONTWL" = "#0081FFFF", "CA_REC" = "#02DA81FF", "ORWA_TWL" = "#80FE1AFF", "ORWA_NONTWL" = "#FDEE02FF", "OR_REC" = "#FFAB00FF", "WA_REC" = "#FF3300FF")

ggplot(data = all_catch, aes(x = YEAR, y = CATCH, fill = ST_FLEET)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  labs(
    x = "Year",
    y = "Catch (mt)",
    fill = ""
  ) +
  ggtitle("Yelloweye Rockfish Commercial Catch (landings + discards)") +
  scale_fill_manual(values = fleetcols) +
  theme_minimal()
```

Catches w/o discard data

```{r}
comm_catch <- catch2 |>
  dplyr::filter(STATE == "CA") |>
  dplyr::rename(CATCH = TOTAL) |>
  rbind(catch_orwa) |>
  dplyr::mutate(ST_FLEET = glue::glue("{STATE}_{FLEET}")) |>
  dplyr::filter(YEAR > 2016) |>
  # dplyr::left_join(discards_all, by = c("YEAR" = "year", "ST_FLEET" = "ST_FLEET")) |>
  dplyr::mutate(
    # discards = dplyr::case_when(is.na(discards) ~ 0,
    #                             TRUE ~ discards),
                TOTAL_CATCH = dplyr::case_when(is.na(CATCH) ~ 0,
                                            TRUE ~ CATCH),
                # TOTAL_CATCH = CATCH + discards,
                SEAS = "1",
                CATCH_SE = 0.01) |>
  dplyr::select(YEAR, SEAS, ST_FLEET, TOTAL_CATCH, CATCH_SE)

all_catch_nodisc <- rbind(hist_c, comm_catch) |>
  dplyr::mutate(
    # YEAR = as.factor(YEAR),
    CATCH = as.numeric(TOTAL_CATCH)
    )

ggplot(data = all_catch_nodisc, aes(x = YEAR, y = CATCH, fill = ST_FLEET)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  labs(
    x = "Year",
    y = "Catch (mt)",
    fill = ""
  ) +
  ggtitle("Yelloweye Rockfish Commercial Catch (landings only)") +
  scale_fill_manual(values = fleetcols) +
  theme_minimal()

```


```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
# catches added for update assessment
ggplot(data = subset(all_catch, YEAR > 2016), aes(x = YEAR, y = TOTAL_CATCH, fill = ST_FLEET)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  labs(
    x = "Year",
    y = "Udpate Yelloweye rockfish catch (mt)",
    fill = ""
  ) +
  scale_fill_manual(values = fleetcols) +
  theme_minimal()

ggplot(data = subset(all_catch_nodisc, YEAR > 2016), aes(x = YEAR, y = TOTAL_CATCH, fill = ST_FLEET)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  labs(
    x = "Year",
    y = "Udpate Yelloweye rockfish catch (mt) (w/o discards)",
    fill = ""
  ) +
  scale_fill_manual(values = fleetcols) +
  theme_minimal()
```

## Notes from class

(Regarding commercial discards)

-   GEMM and WCGOP missing 2004 -\> discards missing then and we will have to make some assumptions to add into landings for that year (values should be the same)

-   use discard rate?

-   2-area model - GEMM might not have the same breaks in the fleets

-   PACFIN YOU can specify

-   WCGOP divided by fleet and won't have to decide by fleet in GEMM

-   do comparisons between the 2

-   look at the sheet for WA recontruction

    -   Fabio is sending

## Double check estimation for update vs assessment

```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
# calculated comm landings
catch_curr_chk <- all_catch |>
  dplyr::mutate(SOURCE = "CURRENT") |>
  dplyr::filter(YEAR > 2001)|>
  dplyr::select(YEAR, ST_FLEET, TOTAL_CATCH, SOURCE)
hist_c_chk <- hist_c |>
  dplyr::filter(YEAR > 2001) |>
  dplyr::mutate(SOURCE = "PREVIOUS") |>
  dplyr::select(YEAR, ST_FLEET, TOTAL_CATCH, SOURCE)
COMB_CHK <- rbind(catch_curr_chk, hist_c_chk)
```

Compare catch from prev assessment to current estimates

```{r}
# ggplot(data = COMB_CHK) +
#   geom_line(aes(x = YEAR, y = as.numeric(TOTAL_CATCH), color = SOURCE), linewidth = 1) +
#   facet_wrap(~ ST_FLEET)
ggplot(data = COMB_CHK, aes(x = YEAR, y=TOTAL_CATCH, fill = SOURCE)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  facet_wrap(~ ST_FLEET, scales = "free_y") +
  labs(
    x="Year",
    y="Catch (mt) with discards"
  )
ggplot(data = COMB_CHK, aes(x = YEAR, y=TOTAL_CATCH, color = SOURCE)) +
  geom_line(linewidth = 1, position=position_jitter(w=0.02, h=0.02)) +
  facet_wrap(~ ST_FLEET, scales = "free_y") +
  labs(
    x="Year",
    y="Catch (mt) with discards"
  )
ggplot() +
  geom_bar(data = subset(COMB_CHK, SOURCE=="PREVIOUS"), aes(x = YEAR, y=TOTAL_CATCH), stat = "identity")+
    geom_line(
    data = subset(COMB_CHK, SOURCE=="CURRENT"), aes(x = YEAR, y=TOTAL_CATCH),
    linewidth = 1, position=position_jitter(w=0.02, h=0.02), color = "red") +
  facet_wrap(~ ST_FLEET, scales = "free_y") +
  labs(
    x="Year",
    y="Catch (mt) with discards"
  )
```

```{r}
#| eval: true
#| echo: false
#| warning: false
#| message: false
# same as above w/o discards
catch_curr_chk2 <- all_catch_nodisc |>
  dplyr::mutate(SOURCE = "CURRENT") |>
  dplyr::filter(YEAR > 2001)|>
  dplyr::select(YEAR, ST_FLEET, TOTAL_CATCH, SOURCE)
hist_c_chk2 <- hist_c |>
  dplyr::filter(YEAR > 2001) |>
  dplyr::mutate(SOURCE = "PREVIOUS") |>
  dplyr::select(YEAR, ST_FLEET, TOTAL_CATCH, SOURCE)
COMB_CHK2 <- rbind(catch_curr_chk2, hist_c_chk2)
ggplot(data = COMB_CHK2, aes(x = YEAR, y=TOTAL_CATCH, fill = SOURCE)) +
  geom_bar(position = "stack", stat = "identity", color = "black") +
  facet_wrap(~ ST_FLEET, scales = "free_y") +
  labs(
    x="Year",
    y="Catch (mt) w/o discards"
  )
ggplot(data = COMB_CHK2, aes(x = YEAR, y=TOTAL_CATCH, color = SOURCE)) +
  geom_line(linewidth = 1, position=position_jitter(w=0.02, h=0.02)) +
  facet_wrap(~ ST_FLEET, scales = "free_y") +
  labs(
    x="Year",
    y="Catch (mt) without discards"
  )

ggplot() +
  geom_bar(data = subset(COMB_CHK2, SOURCE=="PREVIOUS"), aes(x = YEAR, y=TOTAL_CATCH), stat = "identity")+
    geom_line(
    data = subset(COMB_CHK2, SOURCE=="CURRENT"), aes(x = YEAR, y=TOTAL_CATCH),
    linewidth = 1, position=position_jitter(w=0.02, h=0.02), color = "red") +
  facet_wrap(~ ST_FLEET, scales = "free_y") +
  labs(
    x="Year",
    y="Catch (mt) without discards"
  )
```

