---
title: "Comm_age_length_data_forSS"
author: "Juliette Champagnat"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# rm(list=ls())
library(tidyverse)
library(r4ss)

```

## Load data

```{r}

# Get inputs from 2017 assessment that will get carried over to this assessment 
# model_2017_path <- file.path("../../model/2017_yelloweye_model_updated_ss3_exe")
# inputs <- SS_read(dir = model_2017_path, ss_new = TRUE)
ss_data <- SS_readdat(file="../../model/2017_yelloweye_model_updated_ss3_exe/yelloweye_data.ss")
# would using yelloweye_data.ss be different ? with all.equal(ss_data,inputs$dat) they are not exactly equals
# thought summary(ss_data$lencomp==inputs$dat$lencomp) & summary(ss_data$agecomp==inputs$dat$agecomp) -> TRUE


# Get our updated data

## path for data
file_path = "../../Data/processed/length_age_comps/"

## length comps, maal and caal from PacFIN AND WCGOP combined
raw_length_comps_PacFIN_WCGOP <- read.csv(paste0(file_path,'Commercial_length_comps_PacFIN_WCGOP_forSS.csv'))
raw_age_caal_PacFIN_WCGOP <- read.csv(paste0(file_path,'Commercial_caal_PacFIN_WCGOP_forSS.csv'))
raw_age_comps_PacFIN_WCGOP <- read.csv(paste0(file_path,'Commercial_age_comps_PacFIN_WCGOP_forSS.csv'))

## length comps, maal and caal from PacFIN only
raw_length_comps_PacFIN <- read.csv(paste0(file_path,'Commercial_length_comps_PacFIN_forSS.csv'))
raw_age_caal_PacFIN <- read.csv(paste0(file_path,'Commercial_caal_PacFIN_forSS.csv'))
raw_age_comps_PacFIN <- read.csv(paste0(file_path,'Commercial_age_comps_PacFIN_forSS.csv'))

## length comps, maal and caal from PacFIN AND WCGOP combined
raw_length_comps_WCGOP <- read.csv(paste0(file_path,'Commercial_length_comps_WCGOP_forSS.csv'))
raw_age_caal_WCGOP <- read.csv(paste0(file_path,'Commercial_caal_WCGOP_forSS.csv'))
raw_age_comps_WCGOP <- read.csv(paste0(file_path,'Commercial_age_comps_WCGOP_forSS.csv'))


```

## format to good names

```{r}
# length comps
ncol(raw_length_comps_PacFIN_WCGOP);ncol(raw_length_comps_PacFIN);ncol(raw_length_comps_WCGOP);ncol(ss_data$lencomp)
names(raw_length_comps_PacFIN_WCGOP) <- names(raw_length_comps_PacFIN) <- names(raw_length_comps_WCGOP) <- names(ss_data$lencomp)

# caal
ncol(raw_age_caal_PacFIN_WCGOP);ncol(raw_age_caal_PacFIN);ncol(raw_age_caal_WCGOP);ncol(ss_data$agecomp)
names(raw_age_caal_PacFIN_WCGOP) <- names(raw_age_caal_PacFIN) <- names(raw_age_caal_WCGOP) <- names(ss_data$agecomp)

# maal 
ncol(raw_age_comps_PacFIN_WCGOP);ncol(raw_age_comps_PacFIN);ncol(raw_age_comps_WCGOP);ncol(ss_data$agecomp)
names(raw_age_comps_PacFIN_WCGOP) <- names(raw_age_comps_PacFIN) <- names(raw_age_comps_WCGOP) <- names(ss_data$agecomp)

```


## reminder of fleet name

CA.TWL             1
CA.NONTWL          2
ORWA.TWL           4
ORWA.NONTWL        5

# LENGHT

## CA_TWL, PacFIN AND WCGOP

2017 file = 1978-2015
2025 data update = 2016-2024

READY for SS = *YES*

NB: previous assessment was having CA_TWL from PacFIN only because there was so few WCGOP (2005/2008/2010 1 sample each year) that they were not used (and we are still no using them). However now WCGOP has new sample from 2019-2022 and PacFIN from 2019-2024, because the years overlap, length comps now are combined.

```{r}
CA_TWL_PacFIN_WCGOP_length <- ss_data$lencomp %>% filter(fleet==1,year<=2015) %>% 
  bind_rows(raw_length_comps_PacFIN_WCGOP %>% filter(fleet==1,year>2015))
View(CA_TWL_PacFIN_WCGOP_length)

```


## CA_NONTWL, PacFIN

2017 file = 1979 - 2002
2025 data update = no new data from PacFIN

```{r}
raw_length_comps_PacFIN %>% filter(fleet==2,year>2002) #no new data
# check if preivous and updated data matches
raw_length_comps_PacFIN %>% filter(fleet==2,year<=2002) == ss_data$lencomp %>% filter(fleet==2,year<=2002) #yes except for Nsamp
raw_length_comps_PacFIN %>% filter(fleet==2,year<=2002)  %>% pull(Nsamp) - ss_data$lencomp %>% filter(fleet==2,year<=2002) %>% pull(Nsamp) #some non null differences
```

READY for SS = *YES* 

NB: Sample size from 2017 does not matches our current computation for unexplained reason. We still uses the 2017 data.

```{r}
CA_NONTWL_PacFIN_length <- ss_data$lencomp %>% filter(fleet==2,year<=2002) 
View(CA_NONTWL_PacFIN_length)
```

## CA_NONTWL, WCGOP

2017 data = 2004-2015
2025 data update = 2016-2023

```{r}
raw_length_comps_WCGOP %>% 
    filter(fleet==2,year>2002)

#check that previous and updated data are equals
ss_data$lencomp %>% filter(fleet==2,year>2002) == raw_length_comps_WCGOP %>% filter(fleet==2,year>2002,year<2016) # all equals except sample size
ss_data$lencomp %>% filter(fleet==2,year>2002) %>% pull(Nsamp)-raw_length_comps_WCGOP %>% filter(fleet==2,year>2002,year<2016) %>% pull(Nsamp)

```

READY for SS = *YES*

NB: the sample size used in 2017 does not match our values, whereas we have the exact same number of trips and fish samples. We still uses the 2017 data.

```{r}
CA_NONTWL_WCGOP_length <- ss_data$lencomp %>% filter(fleet==2,year>2002) %>% 
  bind_rows(raw_length_comps_WCGOP %>% filter(fleet==2,year>2015))
```

## ORWA_TWL, PacFIN and WCGOP combined

2017 file = 1995-2015 (2016 was included in 2017 but we update it with new data)
2025 data update = 2016-2024

```{r}
# check for overlapping year
ss_data$lencomp %>% filter(fleet==4,year==2015) == raw_length_comps_PacFIN_WCGOP %>%filter(fleet==4,year==2015) 
#2015 is identical-> use data from 2017 datafile
ss_data$lencomp %>% filter(fleet==4,year==2016) == raw_length_comps_PacFIN_WCGOP %>%filter(fleet==4,year==2016) 
# 2016 is different -> use updated data
```

READY for SS = *Yes*

```{r}
ORWA_TWL_PacFIN_WCGOP_length <- ss_data$lencomp %>% filter(fleet==4,year<=2015) %>% 
  bind_rows(raw_length_comps_PacFIN_WCGOP %>% filter(fleet==4,year>2015))
```

## ORWA_NONTWL, PacFIN and WCGOP combined

2017 file = 1980-2015
2025 data update = 2016-2024

```{r}
# check for overlapping year
ss_data$lencomp %>% filter(fleet==5,year==2015) == raw_length_comps_PacFIN_WCGOP %>%filter(fleet==5,year==2015) 
#2015 is identical-> use data from 2017 datafile

```

ready for SS = *YES*

```{r}
ORWA_NONTWL_PacFIN_WCGOP_length <- ss_data$lencomp %>% filter(fleet==5,year<=2015) %>% 
  bind_rows(raw_length_comps_PacFIN_WCGOP %>% filter(fleet==5,year>2015))
```
## Combine everything

ready for SS: *YES*

```{r}

rbind(CA_TWL_PacFIN_WCGOP_length, 
      CA_NONTWL_PacFIN_length,
      CA_NONTWL_WCGOP_length, 
      ORWA_TWL_PacFIN_WCGOP_length,
      ORWA_NONTWL_PacFIN_WCGOP_length)

```



# AGE 

## CA_NONTWL

### CA_NONTWL, CAAL, data from CDFW, Traci file

2017 file = 1978-1988
2025 data update = no new data

```{r}
# check for updated data
raw_age_caal_PacFIN %>% filter(fleet==2,year>1988) 

```

ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.

```{r}
CA_NONTWL_TraciCDFW_caal <- raw_age_caal_PacFIN %>% filter(fleet==2)
View(CA_NONTWL_TraciCDFW_caal)

```


### CA_NONTWL, MAAL ghost, not expanded, data from CDFW, Traci file

2017 file = 1978-1988
2025 data update = no new data

```{r}
# check for updated data
raw_age_comps_PacFIN %>% filter(fleet==-2,year>1988) 
```

ready for SS = *YES*
 
NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.
 
```{r}
CA_NONTWL_TraciCDFW_maal <- raw_age_comps_PacFIN %>% filter(fleet==-2)
View(CA_NONTWL_TraciCDFW_maal)
``` 

### CA_NONTWL, CAAL, WCGOP

2017 file = 2005
2025 data update = no new data

```{r}
# check for updated data
raw_age_caal_WCGOP %>% filter(fleet==2,year>=2005) 
```

Ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.

```{r}
CA_NONTWL_WCGOP_caal <- raw_age_caal_WCGOP %>% filter(fleet==2,year==2005)
```

### CA_NONTWL, WCGOP, ghost MAAL

2017 file = 2015
2025 data update = no new data

```{r}
# check for updated data
raw_age_comps_WCGOP %>% filter(fleet==-2,year>2005) #no new data

# on 2017 data -> caal and maal does not match
ss_data$agecomp %>% filter(fleet==-2,year==2005) %>% select(a0:a65) %>% sum()
ss_data$agecomp %>% filter(fleet==2,year==2005) %>% select(a0:a65) %>% sum()

#not this pb with our updated data
raw_age_comps_WCGOP %>% filter(fleet==-2,year==2005)%>% select(a0:a65) %>% sum()
raw_age_caal_WCGOP %>% filter(fleet==2,year==2005)%>% select(a0:a65) %>% sum()
```
Ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update. Also for this fleet in 2017 file there is a mismatch between caal and maal, not explained.

```{r}
CA_NONTWL_WCGOP_maal <- raw_age_comps_WCGOP %>% filter(fleet==-2,year==2005) 
View(CA_NONTWL_WCGOP_maal)
```

## ORWA_TWL

### ORWA_TWL, CAAL, PacFIN and WCGOP combined

2017 file = 2001-2015 (2016 was included in 2017 but we are updating it with new data)
2025 data update = 2016-2024

```{r}
#check overlapping data
ss_data$agecomp %>% filter(fleet==4,year==2016) 
raw_age_caal_PacFIN_WCGOP %>% filter(fleet==4,year==2016) # new data in 2016 -> use them
```
data ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.

```{r}
ORWA_TWL_PacFIN_WCGOP_caal <- raw_age_caal_PacFIN_WCGOP %>% filter(fleet==4)
View(ORWA_TWL_PacFIN_WCGOP_caal)
```

### ORWA_TWL, ghost MAAL, PacFIN and WCGOP combined

2017 file = 2001-2015 (2016 was included in 2017 but we are updating it with new data)
2025 data update = 2016-2024

```{r}
#check overlapping data
ss_data$agecomp %>% filter(fleet==-4,year==2016) == raw_age_comps_PacFIN_WCGOP %>% filter(fleet==-4,year==2016) # new data in 2016 -> use them
raw_age_comps_PacFIN_WCGOP %>% filter(fleet==-4,year>2016)
```

data ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.

```{r}
ORWA_TWL_PacFIN_WCGOP_maal <-raw_age_comps_PacFIN_WCGOP %>% filter(fleet==-4)
View(ORWA_TWL_PacFIN_WCGOP_maal)

```

## ORWA_NONTWL

### ORWA_NONTWL, CAAL, PacFIN and WCGOP combined

2017 data = 2001-2015
2025 data update = 2016-2024

```{r}
#check for new data
raw_age_caal_PacFIN_WCGOP %>% filter(fleet==5,year>2015) 
```

data ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.

```{r}
ORWA_NONTWL_PacFIN_WCGOP_caal <- raw_age_caal_PacFIN_WCGOP %>% filter(fleet==5)
View(ORWA_NONTWL_PacFIN_WCGOP_caal)

```

### ORWA_NONTWL, ghost MAAL, PacFIN and WCGOP combined

2017 file = 2001-2015 
2025 data update = 2017-2024

```{r}
# new data in 2015 -> use them
raw_age_comps_PacFIN_WCGOP %>% filter(fleet==-5,year>=2016)
```

data ready for SS = *YES*

NB: because of the doubling issue notices in the 2017 caal we're using caal and maal based on our update.

```{r}
ORWA_NONTWL_PacFIN_WCGOP_maal <- raw_age_comps_PacFIN_WCGOP %>% filter(fleet==-5)
```

data ready for SS = *YES*


## Combine everything

TBD when all previous issues are resolved

```{r}

rbind(CA_NONTWL_TraciCDFW_caal,CA_NONTWL_TraciCDFW_maal,
      CA_NONTWL_WCGOP_caal,CA_NONTWL_WCGOP_maal,
      ORWA_TWL_PacFIN_WCGOP_caal, ORWA_TWL_PacFIN_WCGOP_maal,
      ORWA_NONTWL_PacFIN_WCGOP_caal,ORWA_NONTWL_PacFIN_WCGOP_maal
      )
```

