---
title: "Comm_age_length_data_forSS"
author: "Juliette Champagnat"
date: "2025-04-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# rm(list=ls())
library(tidyverse)
# remotes::install_github("pfmc-assessments/pacfintools")
# library(pacfintools)
# remotes::install_github("pfmc-assessments/nwfscSurvey")
# library(nwfscSurvey)
library(r4ss)

```

## Load data

```{r}
#path for data
file_path = "../../Data/processed/length_age_comps/"

# Get inputs from 2017 assessment that will get carried over to this assessment
model_2017_path <- file.path("../../model/2017_yelloweye_model_updated_ss3_exe")
inputs <- SS_read(dir = model_2017_path, ss_new = TRUE)
# wcgop_age_comps

# length comps, maal and caal from PacFIN AND WCGOP
raw_length_comps_comb <- read_csv('../../Data/processed/length_age_comps/_Commercial_length_comps_PacFIN_WCGOP_forSS.csv')
raw_age_caal_comb <- read_csv('../../Data/processed/length_age_comps/Commercial_caal_PacFIN_WCGOP_forSS.csv')
raw_age_comps_comb <- read_csv('../../Data/processed/length_age_comps/_Commercial_age_comps_PacFIN_WCGOP_forSS.csv')

```

## format to good names

```{r}
# length comps
dim(raw_length_comps_comb);dim(ss_data$lencomp)
names(raw_length_comps_comb); names(ss_data$lencomp)
names(raw_length_comps_comb) <- names(ss_data$lencomp)

# caal
dim(raw_age_caal_comb);dim(ss_data$agecomp)
names(raw_age_caal_comb); names(ss_data$agecomp)
names(raw_age_caal_comb) <- names(ss_data$agecomp)

# maal 
dim(raw_age_comps_comb);dim(ss_data$agecomp)
names(raw_age_comps_comb); names(ss_data$agecomp)
names(raw_age_comps_comb) <- names(ss_data$agecomp)
```


## reminder of fleet name

CA.TWL             1
CA.NONTWL          2
ORWA.TWL           4
ORWA.NONTWL        5

# LENGHT

## CA_TWL, PacFIN 

2017 file = 1978-2015
updated data = 2016-2024

READY = no check if we want to mix PacFIN and WCGOP, it's the case for now

```{r}
CA_TWL_PacFIN_length <- ss_data$lencomp %>% filter(fleet==1,year<=2015) %>% 
  bind_rows(raw_length_comps_comb %>% 
  filter(fleet==1,year>2015))
```


## CA_NONTWL, PacFIN

2017 file = 1979 - 2002
update data = 

READY = no check if we want to mix PacFIN and WCGOP, it's the case for now

```{r}
# CA_NONTWL_PacFIN_length <- ss_data$lencomp %>% filter(fleet==2,year<2015) %>% 
#   bind_rows(raw_length_comps_comb %>% 
#   filter(fleet==1,year>2015))

```

## CA_NONTWL, WCGOP

2004-2015

## ORWA_TWL, PacFIN and WCGOP combined

2017 file = 1995-2015
updated data = 2016-2024

```{r}
# check for overlapping year
ss_data$lencomp %>% filter(fleet==4,year==2015) == raw_length_comps_comb %>%filter(fleet==4,year==2015) 
#2015 is identical-> use data from 2017 datafile
ss_data$lencomp %>% filter(fleet==4,year==2016) == raw_length_comps_comb %>%filter(fleet==4,year==2016) 
# 2016 is different -> use updated data
```

READY for SS = Yes

```{r}
ORWA_TWL_PacFIN_WCGOP_length <- ss_data$lencomp %>% filter(fleet==4,year<=2015) %>% 
  bind_rows(raw_length_comps_comb %>% 
  filter(fleet==4,year>2015))
```

## ORWA_NONTWL, PacFIN and WCGOP combined

2017 file = 1980-2015
updated data = 2016-2024

```{r}
# check for overlapping year
ss_data$lencomp %>% filter(fleet==5,year==2015) == raw_length_comps_comb %>%filter(fleet==5,year==2015) 
#2015 is identical-> use data from 2017 datafile

```

ready for SS = YES

```{r}
ORWA_NONTWL_PacFIN_WCGOP_length <- ss_data$lencomp %>% filter(fleet==5,year<=2015) %>% 
  bind_rows(raw_length_comps_comb %>% filter(fleet==5,year>2015))
```
## Combine everything

TBD when all previous issues are resolved

```{r}

```



# AGE 

## CA_NONTWL

### CA_NONTWL, CAAL, data from CDFW, Traci file

2017 file = 1978-1988
no updated data
 *by updated data do you mean nothing new? if there is nothing new, why doesn't it match?*
 Once we get it to match, we can figure out what is going on with that one 7, and then divide by 2, or, just use the new data pull. I think we can justify using the new data if it is the easiest way to fix their mistake.

```{r}
# old data
old_ca_nontwl_caal_1978_1988 <- inputs$dat$agecomp |>
  filter(fleet == 2) |>
  filter(year <= 1988)

# new data
new_ca_nontwl_caal_1978_1988 <- raw_age_caal_comb |>
  filter(fleet == 2) |>
  filter(year <= 1988)

#### Here is the first problem...the data that we filtered does not match. Why?
# Need to go back to filtering method and check this.

# check for updated data
raw_age_caal_comb %>% filter(fleet==2,year>1988) #these are WCGOP data

```

pb of doubling 

```{r}
raw_age_caal_comb %>% filter(fleet==2,year<=1988) %>% nrow()
ss_data$agecomp %>% filter(fleet==2,year<=1988) %>% nrow()

# #another length bin in
# raw_age_caal_comb %>% filter(fleet==2,year==1979) == ss_data$agecomp %>% filter(fleet==2,year==1979) #bin does not match
# 
# raw_age_caal_comb %>% filter(fleet==2,year==1979,Lbin_lo==50)
# ss_data$agecomp %>% filter(fleet==2,year==1979,Lbin_lo==50)
# 
# raw_age_caal_comb %>% filter(fleet==2,year!=1979) == ss_data$agecomp %>% filter(fleet==2,year!=1979) %>% mutate_at(vars(-c('year','month','fleet','sex','part','ageerr','Lbin_lo','Lbin_hi')), .funs=function(x){x/2})

#check for which year /2 caal poses pb -> 1979/2005
ss_data$agecomp %>% filter(fleet==2,year<2016) %>% 
  pivot_longer(cols=-c(year,month,fleet,sex,part,
                                     ageerr,Lbin_lo,Lbin_hi,Nsamp),
                             names_prefix='a',names_to = 'age',
                             values_transform = as.numeric, values_to = "prev_value") %>% 
  mutate(bb=prev_value/2) %>% filter(!bb%in%c(0:100)) %>% View()

```

ready = NO

```{r}
CA_NONTWL_TraciCDFW_caal <- ss_data$agecomp %>% filter(fleet==2,year<=1988) %>% 
  mutate_at(vars(-c('year','month','fleet','sex','part','ageerr','Lbin_lo','Lbin_hi')), .funs=function(x){x/2}) 
  
CA_NONTWL_TraciCDFW_caal %>% View()

```


### CA_NONTWL, MAAL ghost, not expanded, data from CDFW, Traci file

2017 file = 1978-1988
no updated data

*how is it possible that the new MAAL data that you made perfectly matches the old MAAL, but the CAAL does not?*

I would trust the old MAAl data in that the 1979 A24 fish should be a 1, so to fix the doubling problem, we could make that a 2, then divide everything. 

```{r}
# old data
old_ca_nontwl_maal_1978_1988 <- inputs$dat$agecomp |>
  filter(fleet == -2) |>
  filter(year <= 1988)

# let take the old CAAL and mutate it to look like MAAL so we can check the Nsamps for the new data, to help try and figure out what is wrong

old_ca_nontwl_maal_Nsamp <- old_ca_nontwl_caal_1978_1988 %>%
  group_by(year) %>%  # Retain key columns
  summarise(across(starts_with("a"), \(x) sum(x, na.rm = TRUE)),  # Sum age columns
            Nsamp = sum(Nsamp, na.rm = TRUE), .groups = "drop")  # Sum input_n
old_ca_nontwl_maal_Nsamp <- cbind(old_ca_nontwl_maal_1978_1988[,1:8],old_ca_nontwl_maal_Nsamp[,69],old_ca_nontwl_maal_Nsamp[,3:68])

### 1979 seems to be the only problem year, which is consistent with our CAAL problems. The 1979 Age 24 fish should be counted as a 1. In the old MAAL data, it says there is one fish there, where as everything else comparing it to the restructured Nsamp that I made is doubled. so for some reason that one fish didnt get doubled...idk why. 

# new data
new_ca_nontwl_maal_1978_1988 <- raw_age_comps_comb |>
  filter(fleet == -2) |>
  filter(year <= 1988)


# check for updated data
raw_age_comps_comb %>% filter(fleet==-2,year>1988) #these are WCGOP data
```

ready = YES
 
```{r}
CA_NONTWL_TraciCDFW_maal <- ss_data$agecomp %>% filter(fleet==-2,year<=1988) 
``` 

### CA_NONTWL, CAAL, WCGOP

2017 file = 2005
no updated data 

*here it looks like the new data has an extra age. Do you know why?*

```{r}
# check for updated data
raw_age_caal_comb %>% filter(fleet==2,year>=2005) 
```

no doubling issue

```{r}
 ss_data$agecomp %>% filter(fleet==2,year==2005) 
```

ready = YES

```{r}
CA_NONTWL_WCGOP_caal <- ss_data$agecomp %>% filter(fleet==2,year==2005) 
```

### CA_NONTWL, WCGOP, ghost MAAL

2017 file = 2015
no updated data

```{r}
# check for updated data
raw_age_comps_comb %>% filter(fleet==-2,year>2005) #no new data
```

ready = yes 

```{r}
CA_NONTWL_WCGOP_maal <- ss_data$agecomp %>% filter(fleet==-2,year==2005) 
```

## ORWA_TWL

### ORWA_TWL, CAAL, PacFIN and WCGOP combined

2017 file = 2001-2015
updated data = 2016-2024

```{r}
#check overlapping data
ss_data$agecomp %>% filter(fleet==4,year==2016)
raw_age_caal_comb %>% filter(fleet==4,year==2016) # new data in 2016 -> use them
```

doubling problem

```{r}
#check Nsamp -> ok
ss_data$agecomp %>% filter(fleet==4,year<2016) %>% pull(Nsamp) == rowSums(ss_data$agecomp %>% filter(fleet==4,year<2016) %>% select(a0:a65))

#check for which year /2 caal poses pb
ss_data$agecomp %>% filter(fleet==4,year<2016) %>% 
  pivot_longer(cols=-c(year,month,fleet,sex,part,
                                     ageerr,Lbin_lo,Lbin_hi,Nsamp),
                             names_prefix='a',names_to = 'age',
                             values_transform = as.numeric, values_to = "prev_value") %>% 
  mutate(bb=prev_value/2) %>% filter(!bb%in%c(0:100)) %>% View()
# pb year = 2004/2005/2006

# try to remove wcgop data which are 2004-2006
ss_data$agecomp %>% filter(fleet==4,year<2016) %>% 
  #pivot longer ssdata
  pivot_longer(cols=-c(year,month,fleet,sex,part,
                                     Nsamp,ageerr,Lbin_lo,Lbin_hi),
                             names_prefix='a',names_to = 'age',
                             values_transform = as.numeric, values_to = "prev_value") %>% 
  #merge with wcgop
  left_join(wcgop_caal %>% filter(fleet_nb==4,year<2016) %>% 
              dplyr::rename('fleet'=fleet_nb,'part'=partition,'Nsamp'=input_n) %>% 
              pivot_longer(cols=-c(year,month,fleet,sex,part,Nsamp,ageerr,Lbin_lo,Lbin_hi),
                            names_prefix='u',names_to = 'age',
                            values_transform = as.numeric,values_to = 'wcgop_value') %>% select(-Nsamp),
            by=c('year','month','fleet','sex','part','ageerr','Lbin_lo','Lbin_hi','age')) %>% 
  #substract wcgop from initial value
  mutate(value=ifelse(is.na(wcgop_value),prev_value,prev_value-wcgop_value)) %>% # filter(wcgopvalue!=0) %>% 
  #divide by 2 now
  mutate(value=(value/2)) %>% #pull(value) %>% unique()
  #check which year produce weird values -> STOP HERE
  filter(abs(value)==0.5) %>% select(-c(month, sex,part,ageerr))
  # idead was to use this new value to pivot wider and get a good table
    # pivot_wider(id_cols =c(year,month,fleet,sex,part,ageerr,Lbin_lo,Lbin_hi,Nsamp),
    #             values_from = 'value',#values_transform=as.integer,
    #             names_from = 'age',names_prefix = 'a' ) %>% 
  
```


data ready for SS = NO

```{r}
ORWA_TWL_PacFIN_WCGOP_caal <- ss_data$agecomp %>% filter(fleet==4,year<2016) %>% 
  bind_rows(raw_age_caal_comb %>% filter(fleet==4,year>=2016))
# mutate_at(vars(-c('year','month','fleet','sex','part','ageerr','Lbin_lo','Lbin_hi')), .funs=function(x){x/2}) %>% View()
```

### ORWA_TWL, ghost MAAL, PacFIN and WCGOP combined

2017 file = 2001-2016
updated data = 2016-2024

```{r}
#check overlapping data
ss_data$agecomp %>% filter(fleet==-4,year==2016)
raw_age_comps_comb %>% filter(fleet==-4,year==2016) # new data in 2016 -> use them
raw_age_comps_comb %>% filter(fleet==-4,year>2016)
```

data ready for SS = YES

```{r}
ORWA_TWL_PacFIN_WCGOP_maal <- ss_data$agecomp %>% filter(fleet==-4,year<2016) %>% 
  bind_rows(raw_age_comps_comb %>% filter(fleet==-4,year>=2016))
```

## ORWA_NONTWL

### ORWA_NONTWL, CAAL, PacFIN and WCGOP combined

2017 data = 2001-2015
updated data = 2016-2024

```{r}
#check overlapping data
ss_data$agecomp %>% filter(fleet==5,year==2015) == raw_age_caal_comb %>% filter(fleet==5,year==2015) 
# different data in 2015 -> use them ?

raw_age_caal_comb %>% filter(fleet==5,year>2015) 
```


doubling problem

```{r}
#check for which year /2 caal poses pb
ss_data$agecomp %>% filter(fleet==5,year<2016) %>% 
  pivot_longer(cols=-c(year,month,fleet,sex,part,
                                     Nsamp,ageerr,Lbin_lo,Lbin_hi),
                             names_prefix='a',names_to = 'age',
                             values_transform = as.numeric, values_to = "prev_value") %>% 
  mutate(bb=prev_value/2) %>% filter(!bb%in%c(0:10)) %>% select(-c(month,sex,part)) %>% View()

# check if removing wcgop will solve this /2 pb -> 2004/2006/2007 remain weird
ss_data$agecomp %>% filter(fleet==5,year<2016) %>% 
  pivot_longer(cols=-c(year,month,fleet,sex,part,
                                     Nsamp,ageerr,Lbin_lo,Lbin_hi),
                             names_prefix='a',names_to = 'age',
                             values_transform = as.numeric, values_to = "prev_value") %>% 
  left_join(wcgop_caal %>% filter(fleet_nb==5,year<2016) %>% 
              dplyr::rename('fleet'=fleet_nb,'part'=partition,'Nsamp'=input_n) %>% 
              pivot_longer(cols=-c(year,month,fleet,sex,part,Nsamp,ageerr,Lbin_lo,Lbin_hi),
                            names_prefix='u',names_to = 'age',
                            values_transform = as.numeric,values_to = 'wcgop_value') %>% select(-Nsamp),
            by=c('year','month','fleet','sex','part','ageerr','Lbin_lo','Lbin_hi','age')) %>% 
  mutate(value=ifelse(is.na(wcgop_value),prev_value,prev_value-wcgop_value)) %>% # filter(wcgopvalue!=0) %>% 
  mutate(value=(value/2)) %>% #pull(value) %>% unique()
  filter(abs(value)==0.5) %>% select(-c(month, sex,part,ageerr))
    pivot_wider(id_cols =c(year,month,fleet,sex,part,ageerr,Lbin_lo,Lbin_hi,Nsamp),
                values_from = 'value',#values_transform=as.integer,
                names_from = 'age',names_prefix = 'a' ) %>% 
  
```


data ready for SS = NO -> doubling pb not solve

```{r}
ORWA_NONTWL_PacFIN_WCGOP_caal <- ss_data$agecomp %>% filter(fleet==5,year<2016) %>% 
  bind_rows(raw_age_caal_comb %>% filter(fleet==5,year>=2016))
```

### ORWA_NONTWL, ghost MAAL, PacFIN and WCGOP combined

2017 file = 2001-2014 
updated data = 2015-2024

```{r}
#check overlapping data
ss_data$agecomp %>% filter(fleet==-5,year==2015) == raw_age_comps_comb %>% filter(fleet==-5,year==2015) 
# new data in 2015 -> use them
raw_age_comps_comb %>% filter(fleet==-5,year>=2016)
```

data ready for SS = YES

```{r}
ORWA_NONTWL_PacFIN_WCGOP_maal <- ss_data$agecomp %>% filter(fleet==-5,year<2015) %>% 
  bind_rows(raw_age_comps_comb %>% filter(fleet==-5,year>=2015))
```



## Combine everything

TBD when all previous issues are resolved

```{r}

```

