---
title: "rec_lengths"
output: html_document
date: "2025-03-13"
---

```{r setup}
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)
library(r4ss)
library(readr)
library(nwfscSurvey)
library(pacfintools)
```

Load in all data sources and organize by Fleet
1. recfin - lengths only, make sure to get up to date pull after March 20th
2. Julia's CRFS data (CRFS_ages_yeye.xlsx) - ages and lengths
3. MRFSS Bio data - Oregon, from Ali
4. recfin - WA length and age

Fleets:
3 - CA_REC *need to calculate sample sizes
6 - OR_REC *need to calculate sample sizes
7 - WA_REC *need to calculate sample sizes
8 - CA_CPFV (observer program) *need to calculate sample sizes
9 - OR_CPFV (observer program) *need to calculate sample sizes

```{r}
## Load in recfin data and format
recfin_bio <- read.csv("../../Data/raw/confidential/SD501_1983_2024_march.csv")

  # Rename columns to match nwfsc code inputs
names(recfin_bio)[names(recfin_bio) == "RECFIN_YEAR"] <- "year"
names(recfin_bio)[names(recfin_bio) == "SPECIES_NAME"] <- "common_name"
names(recfin_bio)[names(recfin_bio) == "RECFIN_SEX_CODE"] <- "sex"
names(recfin_bio)[names(recfin_bio) == "STATE_NAME"] <- "state"
names(recfin_bio)[names(recfin_bio) == "RECFIN_LENGTH_MM"] <- "length"
recfin_bio$length <- recfin_bio$length/10

## Building Fleet 3 and Fleet 8:
ca_rec <- recfin_bio |> 
  filter(state == "CALIFORNIA") |>
  filter(year != 2020) # Removed the single observation in 2020 (small sample size for that year)

  # Remove CPFV data to make fleets 3 and 8; SORTING BY SOURCE CODE INSTERAD OF MODE NAME.
  # Also, make sure everything is a data.frame!!!
ca_cpfv <- ca_rec |> filter(SOURCE_CODE == "CPFV")
ca_rec <- ca_rec[ca_rec$SOURCE_CODE != "CPFV", ]
ca_rec <- as.data.frame(ca_rec)
ca_cpfv <- as.data.frame(ca_cpfv)

## Building Fleet 6 and Fleet 9:

or_rec <- recfin_bio %>% filter(state == "OREGON")

  # Remove CPFV data to make fleets 3 and 8
  # Also, make sure everything is a data.frame!!!
or_cpfv <- or_rec |> filter(SOURCE_CODE == "CPFV")
or_rec <- or_rec[or_rec$SOURCE_CODE != "CPFV", ]
or_rec <- as.data.frame(or_rec)
or_cpfv <- as.data.frame(or_cpfv)

## Building Fleet 7:

wa_rec <- recfin_bio %>% filter(state == "WASHINGTON")

  # Make sure everything is a data.frame!!!
wa_rec <- as.data.frame(wa_rec)

```

#Data processing: select and filter relevant data, then organize to match format of SS file. Then, load 2017 data and compare with bubble plots.

# For comparison plots - only compare new data from 2004 and on.


Fleet 3 - CA_REC: Recfin, and Don Pearson data (lengths come with the ages)

#Problems: recfin data only goes back to 2004, I bet Don Pearson's data goes back to 1979
#for now, just compare current recfin 2004-2024 to old assessment 2004-2017
#2020 had only one fish
All the data for CA_REC fleet was in RecFin - no outside sources were needed

```{r}
# Organize length data into bins
ca_rec_lengths_list <- get_raw_comps(
  data = ca_rec,
  comp_bins = seq(10, 74, 2),
  comp_column_name = "length",
  input_n_method = "total_samples",
  two_sex_comps = FALSE,
  month = 7,
  fleet = 3,
  partition = 0,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
ca_rec_lengths <- as.data.frame(ca_rec_lengths_list)
colnames(ca_rec_lengths) <- gsub("^unsexed\\.", "", colnames(ca_rec_lengths))
colnames(ca_rec_lengths) <- gsub("^u", "L", colnames(ca_rec_lengths))

# Save
write_csv(ca_rec_lengths, file.path("../../Data/processed/rec_comps/ca_rec_lengths.csv"))

# Comparison plots for 2004-2024 showing that the 2017 CA_REC fleet excludes party/charter boats (CPFV)

  # Read new data
ca_rec_lengths_new <- read.csv("../../Data/processed/rec_comps/ca_rec_lengths.csv") |>
  mutate(assessment = "New_Data") |>
  dplyr::select(year, assessment, L10:L74) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("L", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length),
  )

  # Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
ca_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 3) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

  # Put the long data frames together, filter for all data >= 2004, and bubble plot!
together_rec <- rbind(ca_lengths_old, ca_rec_lengths_new)
together_rec <- together_rec[together_rec$year >= 2004,]
comparison_plot_ca <- together_rec |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("CA_REC Length Comparison")
comparison_plot_ca

  #Save plot
ggsave(plot = comparison_plot_ca, "ca_rec_length_comp_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

### Confirming that CA_REC fleet does not include CPFV - next compare current recfin cpfv data to 2017

# Next step is to combine all 2017 CA_REC data with New 2017-2024 data

```

Fleet 6 - OR_REC: MRFSS and ORBS combined, 
plus lengths with WDFW ages (1979-2002) - these should be included already in recfin
ODFW ages (2009-2016/present?) *Ali has this and will upload* lengths should also be in recfin...
Make sure to separate out 

*not included in recfin, see Ali's uploaded data*
Recfin includes data starting in 2001
MRFSS data goes from 1980 to 2003 - *roll over from 2017* For benchmark, use only measured lengths
ORBS data? *just start in 2004 to avoid the overlap of mrfss and orbs data*
WDFW/ODFW ages?

```{r}
# Organize length data into bins
or_rec_lengths_list <- get_raw_comps(
  data = or_rec,
  comp_bins = seq(10, 74, 2),
  comp_column_name = "length",
  input_n_method = "total_samples",
  two_sex_comps = FALSE,
  month = 7,
  fleet = 6,
  partition = 0,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
or_rec_lengths <- as.data.frame(or_rec_lengths_list)
colnames(or_rec_lengths) <- gsub("^unsexed\\.", "", colnames(or_rec_lengths))
colnames(or_rec_lengths) <- gsub("^u", "L", colnames(or_rec_lengths))
# Filter out everything before 2004 to avoid MRFSS data
or_rec_lengths <- or_rec_lengths[or_rec_lengths$year>=2004,]

# Save
write_csv(or_rec_lengths, file.path("../../Data/processed/rec_comps/or_rec_lengths.csv"))

# Comparison plots for 2004-2024 showing that the 2017 OR_REC fleet excludes party/charter boats (CPFV)

  # Read new data
or_rec_lengths_new <- read.csv("../../Data/processed/rec_comps/or_rec_lengths.csv") |>
  mutate(assessment = "New_Data") |>
  dplyr::select(year, assessment, L10:L74) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("L", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length),
  )

  # Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
or_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 6) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

  # Put the long data frames together, filter for all data >= 2004, and bubble plot!
together_rec <- rbind(or_lengths_old, or_rec_lengths_new)
together_rec <- together_rec[together_rec$year >= 2004,]
comparison_plot_or <- together_rec |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("OR_REC Length Comparison")
comparison_plot_or

  #Save plot
ggsave(plot = comparison_plot_or, "or_rec_length_comp_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

### Confirming that OR_REC fleet does not include CPFV - next compare current recfin cpfv data to 2017

# Next step is to combine all 2017 OR_REC data with New 2017-2024 data
```

WA_REC: WDFW data (starting in 1981)

Super weird...the RecFin data starts in 1983, the very old dates dont line up at all. (double check!! I bet some of those old lengths may have come from Age data that we havent gotten yet??)
Is it possible to get this data from somewhere else?
*the assessment document said they got their data directly from WDFW, not RecFin*
use previous assessment and ignore old recfin data? ask higher powers

```{r}
# Organize length data into bins
wa_rec_lengths_list <- get_raw_comps(
  data = wa_rec,
  comp_bins = seq(10, 74, 2),
  comp_column_name = "length",
  input_n_method = "total_samples",
  two_sex_comps = FALSE,
  month = 7,
  fleet = 7,
  partition = 0,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
wa_rec_lengths <- as.data.frame(wa_rec_lengths_list)
colnames(wa_rec_lengths) <- gsub("^unsexed\\.", "", colnames(wa_rec_lengths))
colnames(wa_rec_lengths) <- gsub("^u", "L", colnames(wa_rec_lengths))

# Save
write_csv(wa_rec_lengths, file.path("../../Data/processed/rec_comps/wa_rec_lengths.csv"))

# Comparison plots for 2004-2024 showing that the 2017 WA_REC fleet excludes party/charter boats (CPFV)

  # Read new data
wa_rec_lengths_new <- read.csv("../../Data/processed/rec_comps/wa_rec_lengths.csv") |>
  mutate(assessment = "New_Data") |>
  dplyr::select(year, assessment, L10:L74) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("L", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length),
  )

  # Read old data
inputs_old <- r4ss::SS_read(dir=file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
wa_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 7) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

  # Put the long data frames together, filter for all data >= 2004, and bubble plot!
together_rec <- rbind(wa_lengths_old, wa_rec_lengths_new)
#together_rec <- together_rec[together_rec$year >= 1983,]
comparison_plot_ca <- together_rec |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("WA_REC Length Comparison")
comparison_plot_ca

  #Save plot
ggsave(plot = comparison_plot_ca, "wa_rec_length_comp_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

## It looks like RecFin has a lot of data that wasn't in the assessment, but is also missing all the really old data that they used (they got it directly from WDFW)

# Next step is to combine all 2017 CA_REC data with New 2017-2024 data
```

Fleet 8 - CA_Observer: CPFV, Melissa Monk, starting in 1987-1999, 2004-2016

*ask Julia about CRFS data, same thing? But that only goes from 2009 to 2016*
check the catch data, they used CRFS starting in 2004

```{r}
ca_cpfv

# Organize length data into bins
ca_obs_lengths_list <- get_raw_comps(
  data = ca_cpfv,
  comp_bins = seq(10, 74, 2),
  comp_column_name = "length",
  input_n_method = "total_samples",
  two_sex_comps = FALSE,
  month = 7,
  fleet = 8,
  partition = 0,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
ca_obs_lengths <- as.data.frame(ca_obs_lengths_list)
colnames(ca_obs_lengths) <- gsub("^unsexed\\.", "", colnames(ca_obs_lengths))
colnames(ca_obs_lengths) <- gsub("^u", "L", colnames(ca_obs_lengths))

# Save
write_csv(ca_obs_lengths, file.path("../../Data/processed/rec_comps/ca_obs_lengths.csv"))

# Comparison plots for 2004-2024 

  # Read new data
ca_obs_lengths_new <- read.csv("../../Data/processed/rec_comps/ca_obs_lengths.csv") |>
  mutate(assessment = "New_Data") |>
  dplyr::select(year, assessment, L10:L74) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("L", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length),
  )

  # Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
ca_obs_lengths_old <- inputs_old$dat$lencomp
ca_obs_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 8) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

  # Put the long data frames together, filter for all data >= 2004, and bubble plot!
together_obs <- rbind(ca_obs_lengths_old, ca_obs_lengths_new)
together_obs <- together_obs[together_obs$year >= 2004,]
comparison_plot_ca <- together_obs |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("CA_observer Length Comparison")
comparison_plot_ca

  #Save plot
ggsave(plot = comparison_plot_ca, "ca_obs_length_comp_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

# Next step is to combine all 2017 CA_obs data with New 2017-2024 data

```

OR_Observer: ODFW data, starting in 2003 (lets start in 2004 again, to avoid mrfss data)

```{r}
#or_cpfv

# Organize length data into bins
or_obs_lengths_list <- get_raw_comps(
  data = or_cpfv,
  comp_bins = seq(10, 74, 2),
  comp_column_name = "length",
  input_n_method = "total_samples",
  two_sex_comps = FALSE,
  month = 7,
  fleet = 9,
  partition = 0,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
or_obs_lengths <- as.data.frame(or_obs_lengths_list)
colnames(or_obs_lengths) <- gsub("^unsexed\\.", "", colnames(or_obs_lengths))
colnames(or_obs_lengths) <- gsub("^u", "L", colnames(or_obs_lengths))

# Save
write_csv(or_obs_lengths, file.path("../../Data/processed/rec_comps/or_obs_lengths.csv"))

# Comparison plots for 2003-2024 

  # Read new data
or_obs_lengths_new <- read.csv("../../Data/processed/rec_comps/or_obs_lengths.csv") |>
  mutate(assessment = "New_Data") |>
  dplyr::select(year, assessment, L10:L74) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("L", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length),
  )

  # Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
or_obs_lengths_old <- inputs_old$dat$lencomp
or_obs_lengths_old <- inputs_old$dat$lencomp |>
  dplyr::filter(fleet == 9) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `l10`:`l74`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "length", values_to = "freq") |>
  dplyr::mutate(length = gsub("^l", "", length)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    length = as.numeric(length)
  )

  # Put the long data frames together, filter for all data >= 2004, and bubble plot!
together_obs <- rbind(or_obs_lengths_old, or_obs_lengths_new)
comparison_plot_or <- together_obs |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = length, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("OR_observer Length Comparison")
comparison_plot_or

  #Save plot
ggsave(plot = comparison_plot_or, "or_obs_length_comp_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

# Next step is to combine all 2017 OR_obs data with New 2017-2024 data
```


#########
Playing with Nsamps predictions

```{r}
# Get inputs from 2017 assessment that will get carried over to this assessment
model_2017_path <- file.path("../../model/2017_yelloweye_model_updated_ss3_exe")
inputs <- SS_read(dir = model_2017_path, ss_new = TRUE)

# CA_REC comparison and lm line
ca_rec_len <- inputs$dat$lencomp  |>
  filter(fleet == 3) |>
  filter(year >= 2004)

new_ca_rec_len <- read_csv(file.path("../../Data/processed/rec_comps/ca_rec_lengths.csv"))
new_ca_rec_len <- new_ca_rec_len |>
  filter(year <= 2016)

compare_input_ca <- as.data.frame(cbind(ca_rec_len$Nsamp,new_ca_rec_len$input_n))
model_ca <- lm(compare_input_ca$V1 ~ compare_input_ca$V2)
model_ca
ggplot(compare_input_ca, aes(V2,V1))+
  geom_point() +
  geom_smooth(method = lm, se = TRUE, col = 'red')

# OR_REC comparison and lm line
or_rec_len <- inputs$dat$lencomp  |>
  filter(fleet == 6) |>
  filter(year >= 2004)

new_or_rec_len <- read_csv(file.path("../../Data/processed/rec_comps/or_rec_lengths.csv"))
new_or_rec_len <- new_or_rec_len |>
  filter(year <= 2016)

compare_input_or <- as.data.frame(cbind(or_rec_len$Nsamp,new_or_rec_len$input_n))
model_or <- lm(compare_input_or$V1 ~ compare_input_or$V2)
model_or
ggplot(compare_input_or, aes(V2,V1))+
  geom_point() +
  geom_smooth(method = lm, se = TRUE, col = 'orange')

# WA_REC comparison and lm line
wa_rec_len <- inputs$dat$lencomp  |>
  filter(fleet == 7) |>
  filter(year >= 1998)

new_wa_rec_len <- read_csv(file.path("../../Data/processed/rec_comps/wa_rec_lengths.csv"))
new_wa_rec_len <- new_wa_rec_len |>
  filter(year <= 2016 & year >= 1998)

compare_input_wa <- as.data.frame(cbind(wa_rec_len$Nsamp,new_wa_rec_len$input_n))
model_wa <- lm(compare_input_wa$V1 ~ compare_input_wa$V2)
model_wa
ggplot(compare_input_wa, aes(V2,V1))+
  geom_point() +
  geom_smooth(method = lm, se = TRUE, col = 'darkgreen')

# CA_OBS comparision lm line
ca_obs_len <- inputs$dat$lencomp  |>
  filter(fleet == 8)|>
  filter(year >= 2008)

new_ca_obs_len <- read_csv(file.path("../../Data/processed/rec_comps/ca_obs_lengths.csv"))
new_ca_obs_len <- new_ca_obs_len |>
  filter(year <= 2016 & year >= 2008)

compare_input_ca_obs <- as.data.frame(cbind(ca_obs_len$Nsamp,new_ca_obs_len$input_n))
model_ca_obs <- lm(compare_input_ca_obs$V1 ~ compare_input_ca_obs$V2)
model_ca_obs
ggplot(compare_input_ca_obs, aes(V2,V1))+
  geom_point() +
  geom_smooth(method = lm, se = TRUE, col = 'blue')


```




