---
title: "rec_lengthcomps_for_datafile"
output: html_document
date: "2025-04-04"
---

Use this document as a source for Rec length comp code for pulling data into each of the model runs. I will comment out the code so that it is easy to see what should go into the models.

```{r setup, include=FALSE}
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)
library(r4ss)
library(readr)
library(nwfscSurvey)
library(pacfintools)
```

Read in 2017 model data
```{r}
# Get inputs from 2017 assessment that will get carried over to this assessment
model_2017_path <- file.path("../../model/2017_yelloweye_model_updated_ss3_exe")
inputs <- SS_read(dir = model_2017_path, ss_new = TRUE)
```

1. CA_REC, RecFin and Don's data, 1979-2024
  Remember, we filtered out 2020 data because there was only 1 length available.
  There was a perfect match for data, so no updates for old data. 
  Just add the new stuff on top, with fixed Nsamps
```{r}
# Use this first section to grab the old data with out any changes!
ca_rec_len <- inputs$dat$lencomp  |>
  filter(fleet == 3)

## For the Update, add this next section to add data from 2017-2024
new_ca_rec_len <- read_csv(file.path("../../Data/processed/rec_comps/ca_rec_lengths.csv"))
names(new_ca_rec_len) <- names(ca_rec_len) # Replace col names so they match everything else
new_ca_rec_len <- new_ca_rec_len |>
  filter(year >= 2017) # only grab newest years
  ### Calculate Nsamps using linear equation y = 4.6 + 0.732x
x <- new_ca_rec_len$Nsamp
y <-  4.6 + 0.732*x
new_ca_rec_len$Nsamp <- y

#put them together
ca_rec_len <- rbind(ca_rec_len,new_ca_rec_len) 
```

2. OR_REC, MRFFS and ORBS combined, 1979-2024
  There was a *close* match for data, so no updates for old data. 
  Might be worth investigating these data sources for the future assessment.
  Just add the new stuff on top, with fixed Nsamps
```{r}
# Use this first section to grab the old data with out any changes!
or_rec_len <- inputs$dat$lencomp  |>
  filter(fleet == 6)

## For the Update, add this next section to add data from 2017-2024
new_or_rec_len <- read_csv(file.path("../../Data/processed/rec_comps/or_rec_lengths.csv"))
names(new_or_rec_len) <- names(or_rec_len) # Replace col names so they match everything else
new_or_rec_len <- new_or_rec_len |>
  filter(year >= 2017) # only grab newest years
  ### Calculate Nsamps using linear equation y = 0.341 + 1.1x
x <- new_or_rec_len$Nsamp
y <-  0.341 + 1.1*x
new_or_rec_len$Nsamp <- y

#put them together
or_rec_len <- rbind(or_rec_len,new_or_rec_len) 
```

3. WA_REC, 1981, 1987, 1996-2024
  There was a *close* match for data, no updates for old data. 
  Might be worth investigating these data sources for the future assessment.
  Just add the new stuff on top, with fixed Nsamps
```{r}
# Use this first section to grab the old data with out any changes!
wa_rec_len <- inputs$dat$lencomp  |>
  filter(fleet == 7)

## For the Update, add this next section to add data from 2017-2024
new_wa_rec_len <- read_csv(file.path("../../Data/processed/rec_comps/wa_rec_lengths.csv"))
names(new_wa_rec_len) <- names(wa_rec_len) # Replace col names so they match everything else
new_wa_rec_len <- new_wa_rec_len |>
  filter(year >= 2017) # only grab newest years, no data for 2016
  ### Calculate Nsamps using linear equation y = 2.02 + 0.17x
x <- new_wa_rec_len$Nsamp
y <-  2.02 + 0.17*x
new_wa_rec_len$Nsamp <- y

#put them together
wa_rec_len <- rbind(wa_rec_len,new_wa_rec_len) 
```

4. CA_OBS, 1987-2024
  There was a *close* match for data, no updates for old data. 
  Might be worth investigating these data sources for the future assessment.
  Just add the new stuff on top, with fixed Nsamps
```{r}
# Use this first section to grab the old data with out any changes!
ca_obs_len <- inputs$dat$lencomp  |>
  filter(fleet == 8)

## For the Update, add this next section to add data from 2017-2024
new_ca_obs_len <- read_csv(file.path("../../Data/processed/rec_comps/ca_obs_lengths.csv"))
names(new_ca_obs_len) <- names(ca_obs_len) # Replace col names so they match everything else
new_ca_obs_len <- new_ca_obs_len |>
  filter(year >= 2017) # only grab newest years
  ### Calculate Nsamps using linear equation y = 1.48 + 0.73x
x <- new_ca_obs_len$Nsamp
y <-  1.48 + 0.73*x
new_ca_obs_len$Nsamp <- y

#put them together
ca_obs_len <- rbind(ca_obs_len,new_ca_obs_len) 
```

5. OR_OBS, 2003-2024
  There was a perfect match for length data. 
  However, the Nsamps in the 2017 assessment are whole numbers.
  For some reason they were rounded. Vlada found the raw Nsamp numbers. So for the update, we will replace them.
  Also fix the Nsamps using linear equation
```{r}
# Use this first section to grab the old data with out any changes!
or_obs_len <- inputs$dat$lencomp  |>
  filter(fleet == 9)

# Fix Nsamps so they aren't rounded numbers using Vlada's data
or_obs_len$Nsamp <- c(2.276,13.898,15.312,30.348,30.176,29.142,18.416,14.76,15.14,41.42,26.658,34.626,18.002,19.864)

## For the Update, use only the new data, so that it has both updated years and updated Nsamps 
new_or_obs_len <- read_csv(file.path("../../Data/processed/rec_comps/or_obs_lengths.csv"))
names(new_or_obs_len) <- names(or_obs_len) # Replace col names so they match everything else
new_or_obs_len <- new_or_obs_len |>
  filter(year >= 2017) # only grab newest years
  ### Calculate Nsamps using linear equation y = 5.76 + 0.415x
x <- new_or_obs_len$Nsamp
y <-  5.76 + 0.415*x
new_or_obs_len$Nsamp <- y

#put them together
or_obs_len <- rbind(or_obs_len,new_or_obs_len) 
```