---
title: "rec_ages_for_datafile"
output: html_document
date: "2025-03-30"
---

```{r setup, include=FALSE}
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)
library(r4ss)
library(readr)
library(nwfscSurvey)
library(pacfintools)
```

Read in 2017 model data
```{r}
# Get inputs from 2017 assessment that will get carried over to this assessment
model_2017_path <- file.path("../../model/2017_yelloweye_model_updated_ss3_exe")
inputs <- SS_read(dir = model_2017_path, ss_new = TRUE)
```

1. CA_REC CAAL 1983-1996, aged by WDFW
  *no change* pull in old data

```{r}
ca_rec_caal_1983_1996 <- inputs$dat$agecomp |>
  filter(fleet == 3) |>
  filter(year == 1983 | year == 1996) |>
  filter(ageerr == 1)
```

2. CA_REC MAAL Ghost 1983-1996, aged by WDFW
  *no change* pull in old data

```{r}
ca_rec_maal_1983_1996 <- inputs$dat$agecomp |>
  filter(fleet == -3) |>
  filter(year == 1983 | year == 1996) |>
  filter(ageerr == 1)
```

3. CA_REC CAAL Don Pearson Data 1979-1984, aged by Betty
  *divide all Nsamp and Ages by 2* - everything was doubled by accident in last assessment
```{r}
ca_rec_caal_Don <- inputs$dat$agecomp |>
  filter(fleet == 3) |>
  filter(year >= 1979 & year <= 1984) |>
  filter(ageerr == 2)
ca_rec_caal_Don[,9:75] <- ca_rec_caal_Don[,9:75]/2
```

4. CA_REC CAAL John Budrick Data 2009-2016, aged by Betty
  *divide all Nsamp and Ages by 2* - No new data, all data we got matched perfectly with 2017 assessment data, so just use old data.
```{r}
ca_rec_caal_John <- inputs$dat$agecomp |>
  filter(fleet == 3) |>
  filter(year >= 2009 & year <= 2016)
ca_rec_caal_John[,9:75] <- ca_rec_caal_John[,9:75]/2
```

5. CA_REC MAAL Ghost, Don's data
  *need to rebuild from CAAL* - Nsamp column was duplicated (and is wrong), so it shifted all of the ages forward by 1
```{r}
ca_rec_maal_Don <- inputs$dat$agecomp |>
  filter(fleet == -3) |>
  filter(year >= 1979 & year <= 1984) |>
  filter(ageerr == 2)
```

```{r}


```

```{r}


```

Finish and put it all together 
```{r}
mod$dat$agecomp <- dplyr::bind_rows(...) # list all the dfs that I just made

```
