---
title: "rec_ages"
output: html_document
date: "2025-03-13"
---

```{r setup}
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)
library(r4ss)
library(readr)
library(nwfscSurvey)
library(pacfintools)
```

*if there are no new CA age data since 2016, then dont worry about it*

Load in all data sources
1. CA ages from Julia: "CRFS_ages_yeye" * no updates
2. OR ages from Ali: "OR_non_standard_yelloweye_2009-2016_nonconfid" * no updates
3. WA ages from Fabio * updated in April

Fleets:
3 - CA_REC 
6 - OR_REC 
7 - WA_REC 

```{r}
# Load CA ages and format
ca_ages_raw <- read_csv("../../Data/raw/nonconfidential/CRFS_ages_yeye.csv")

 # Rename columns to match nwfsc code inputs
names(ca_ages_raw)[names(ca_ages_raw) == "Year"] <- "year"
names(ca_ages_raw)[names(ca_ages_raw) == "Final Age"] <- "age"
names(ca_ages_raw)[names(ca_ages_raw) == "Sex"] <- "sex"
  # Change the sex column to a uniform number
ca_ages_raw[,17] = "U"
  # One length had only total length, so I used the Assessment document's equation to get a new value (0.9824*365+0.7522 = 359.3282
ca_ages_raw[1,13] = 359.3282
names(ca_ages_raw)[names(ca_ages_raw) == "Fork Length (mm)"] <- "length"
ca_ages_raw$length <- ca_ages_raw$length/10
ca_ages_raw <- as.data.frame(ca_ages_raw)

# Load OR ages and format
or_ages_raw <- read_csv("../../Data/raw/nonconfidential/OR_non_standard_yelloweye_2009-2016_nonconfid.csv")

  # Rename columns to match nwfsc code inputs
names(or_ages_raw)[names(or_ages_raw) == "YYYY"] <- "year"
names(or_ages_raw)[names(or_ages_raw) == "Age"] <- "age"
names(or_ages_raw)[names(or_ages_raw) == "Sex"] <- "sex"
  # Change the sex column to a uniform number
or_ages_raw[,37] = "U"
names(or_ages_raw)[names(or_ages_raw) == "SampleLength"] <- "length"
or_ages_raw$length <- or_ages_raw$length/10
or_ages_raw <- as.data.frame(or_ages_raw)

# Load WA ages and format
wa_ages_raw <- read_csv("../../Data/raw/nonconfidential/wa_rec_ages_April4.csv")
  
  # Rename columns 
names(wa_ages_raw)[names(wa_ages_raw) == "sample_year"] <- "year"
names(wa_ages_raw)[names(wa_ages_raw) == "species_name"] <- "common_name"
names(wa_ages_raw)[names(wa_ages_raw) == "sex_name"] <- "sex"
names(wa_ages_raw)[names(wa_ages_raw) == "best_age"] <- "age"
names(wa_ages_raw)[names(wa_ages_raw) == "fish_length_cm"] <- "length"
  # Change the sex column to U and fix length
wa_ages_raw[,18] = "U"
wa_ages_raw <- as.data.frame(wa_ages_raw)
wa_ages_raw <- wa_ages_raw |> filter(!is.na(age))

```

#Data processing: select and filter relevant data, then organize to match format of SS file. Then, load 2017 data and compare with bubble plots.


CA_REC: CAAL, CDFW. We only need to find CDFW data starting in 2009 (John Budrick). Age Error 2

*this data should pretty much exactly match, except we might have an Nsamps problem?*

```{r}
# Organize data into age bins conditional on length bins. Make sure it's a single sex.
ca_rec_caal <- get_raw_caal(
  data = ca_ages_raw,
  len_bins = seq(10, 74, 2),
  age_bins = seq(0, 65, 1),
  length_column_name = "length",
  age_column_name = "age",
  month = 7,
  fleet = 3,
  partition = 0,
  ageerr = 2,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data fram
ca_rec_caal <- data.frame(ca_rec_caal)
colnames(ca_rec_caal) <- gsub("^u", "A", colnames(ca_rec_caal))

# Save
write.csv(ca_rec_caal, file.path("../../Data/processed/rec_comps/ca_rec_caal.csv"))

# Comparison plots for 2009-2016
  # Read new data
ca_rec_caal_new <- read.csv("../../Data/processed/rec_comps/ca_rec_caal.csv") |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, A0:A65) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("A", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age),
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
ca_rec_caal_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == 3) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age)
  )

  # Put the long data frames together, filter for all data >= 2009, and bubble plot!
together_caal <- rbind(ca_rec_caal_old, ca_rec_caal_new)
together_caal <- together_caal[together_caal$year >= 2009,]
comparison_plot_ca <- together_caal |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("CA_REC CAAL Comparison")
comparison_plot_ca

  #Save plot
ggsave(plot = comparison_plot_ca, "ca_rec_caal_comparisons.png", path = file.path("../../Rcode/length_age_comps"))
```

CA_REC: MAAL, CDFW. We only need to find CDFW data starting in 2009 (John Budrick). Age Error 2

```{r}
# Organize data into age bins. Make sure it's a single sex.
ca_rec_maal <- get_raw_comps(
  data = ca_ages_raw,
  comp_bins = seq(0, 65, 1),
  comp_column_name = "age",
  input_n_method = c("total_samples"),
  two_sex_comps = FALSE,
  month = 7,
  fleet = -3,
  partition = 0,
  ageerr = 2,
  Lbin_lo = -1,
  Lbin_hi = -1,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
ca_rec_maal <- data.frame(ca_rec_maal)
colnames(ca_rec_maal) <- gsub("^unsexed\\.", "", colnames(ca_rec_maal))
colnames(ca_rec_maal) <- gsub("^u", "A", colnames(ca_rec_maal))

# Save
write.csv(ca_rec_maal, file.path("../../Data/processed/rec_comps/ca_rec_maal.csv"))

# Comparison plots for 2009-2016
  # Read new data
ca_rec_maal_new <- read.csv("../../Data/processed/rec_comps/ca_rec_maal.csv") |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, A0:A65) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("A", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age),
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
ca_rec_maal_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == -3) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age)
  )

  # Put the long data frames together and bubble plot!
together <- rbind(ca_rec_maal_old, ca_rec_maal_new)
together <- together[together$year >= 2009,]
comparison_plot <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("CA_REC MAAL Comparison")
comparison_plot
  
  #Save plot
ggsave(plot = comparison_plot, "ca_rec_maal_comparisons.png", path = file.path("../../Rcode/length_age_comps"))
```

OR_REC: CAAL, WDFW ager data. Starting in 1979. Age Error 1 AND 2????

For OR ages, lets go with Age Error 1 (WDFW) for now...double check with Ali, but in the raw document, the LabID = 2, maybe meaning it was aged with NWFSC? Or did Ali say WDFW ages all their fish?

```{r}
# Organize data into age bins conditional on length bins. Make sure it's a single sex.
or_rec_caal <- get_raw_caal(
  data = or_ages_raw,
  len_bins = seq(10, 74, 2),
  age_bins = seq(0, 65, 1),
  length_column_name = "length",
  age_column_name = "age",
  month = 7,
  fleet = 6,
  partition = 0,
  ageerr = 2,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data fram
or_rec_caal <- data.frame(or_rec_caal)
colnames(or_rec_caal) <- gsub("^u", "A", colnames(or_rec_caal))

# Save
write_csv(or_rec_caal, file.path("../../Data/processed/rec_comps/or_rec_caal.csv"))

# Comparison plots for 2009-2016
  # Read new data
or_rec_caal_new <- read.csv("../../Data/processed/rec_comps/or_rec_caal.csv") |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, A0:A65) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("A", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age),
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
or_rec_caal_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == 6) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age)
  )

  # Put the long data frames together, filter for all data >= 2009, and bubble plot!
together_caal <- rbind(or_rec_caal_old, or_rec_caal_new)
together_caal <- together_caal[together_caal$year >= 2009,]
comparison_plot_ca <- together_caal |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("OR_REC CAAL Comparison")
comparison_plot_ca

  #Save plot
ggsave(plot = comparison_plot_ca, "or_rec_caal_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

# there are two more ages present (31 and 17) I wonder why these were not included. It looks like they were aged all in the same way as the other ages for that year. New data?
```

OR_REC: MAAL, WDFW ager data. Starting in 1979. Age Error 1 AND 2

```{r}
# Organize data into age bins. Make sure it's a single sex.
or_rec_maal <- get_raw_comps(
  data = or_ages_raw,
  comp_bins = seq(0, 65, 1),
  comp_column_name = "age",
  input_n_method = c("total_samples"),
  two_sex_comps = FALSE,
  month = 7,
  fleet = -6,
  partition = 0,
  ageerr = 2,
  Lbin_lo = -1,
  Lbin_hi = -1,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
or_rec_maal <- data.frame(or_rec_maal)
colnames(or_rec_maal) <- gsub("^unsexed\\.", "", colnames(or_rec_maal))
colnames(or_rec_maal) <- gsub("^u", "A", colnames(or_rec_maal))

# Save
write.csv(or_rec_maal, file.path("../../Data/processed/rec_comps/or_rec_maal.csv"))

# Comparison plots for 2009-2016
  # Read new data
or_rec_maal_new <- read.csv("../../Data/processed/rec_comps/or_rec_maal.csv") |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, A0:A65) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("A", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age),
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
or_rec_maal_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == -6) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age)
  )

  # Put the long data frames together and bubble plot!
together <- rbind(or_rec_maal_old, or_rec_maal_new)
together <- together[together$year >= 2009,]
comparison_plot <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("OR_REC MAAL Comparison")
comparison_plot
  
  #Save plot
ggsave(plot = comparison_plot, "or_rec_maal_comparisons.png", path = file.path("../../Rcode/length_age_comps"))
```

WA_REC: CAAL, WDFW ager data. Starting in 1998. includes MRFSS and OSP. Age Error 1

```{r}
# Organize data into age bins conditional on length bins. Make sure it's a single sex.
wa_rec_caal <- get_raw_caal(
  data = wa_ages_raw,
  len_bins = seq(10, 74, 2),
  age_bins = seq(0, 65, 1),
  length_column_name = "length",
  age_column_name = "age",
  month = 7,
  fleet = 7,
  partition = 0,
  ageerr = 1,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
wa_rec_caal <- data.frame(wa_rec_caal)
colnames(wa_rec_caal) <- gsub("^u", "A", colnames(wa_rec_caal))

# Save
write_csv(wa_rec_caal, file.path("../../Data/processed/rec_comps/wa_rec_caal.csv"))

# Comparison plots for 1998-2024
  # Read new data
wa_rec_caal_new <- read_csv("../../Data/processed/rec_comps/wa_rec_caal.csv") |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, A0:A65) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("A", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age),
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
wa_rec_caal_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == 7) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age)
  )

  # Put the long data frames together and bubble plot!
together <- rbind(wa_rec_caal_old, wa_rec_caal_new)
comparison_plot_wa <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("WA_REC CAAL Comparison")
comparison_plot_wa
  
  #Save plot
ggsave(plot = comparison_plot_wa, "wa_rec_caal_comparisons.png", path = file.path("../../Rcode/length_age_comps"))

```

WA_REC: MAAL, WDFW ager data. Starting in 1998. includes MRFSS and OSP
Some ages have changed 2004-2015 from previous assessment to our data

```{r}
# Organize data into age bins. Make sure it's a single sex.
wa_rec_maal <- get_raw_comps(
  data = wa_ages_raw,
  comp_bins = seq(0, 65, 1),
  comp_column_name = "age",
  input_n_method = "total_samples",
  two_sex_comps = FALSE,
  month = 7,
  fleet = -7,
  partition = 0,
  ageerr = 1,
  Lbin_lo = -1,
  Lbin_hi = -1,
  dir = NULL,
  printfolder = "",
  verbose = TRUE
)

# Make it a readable data frame
wa_rec_maal <- data.frame(wa_rec_maal)
colnames(wa_rec_maal) <- gsub("^unsexed\\.", "", colnames(wa_rec_maal))
colnames(wa_rec_maal) <- gsub("^u", "A", colnames(wa_rec_maal))

# Save
write_csv(wa_rec_maal, file.path("../../Data/processed/rec_comps/wa_rec_maal.csv"))

# Comparison plots for 1998-2024
  # Read new data
wa_rec_maal_new <- read_csv("../../Data/processed/rec_comps/wa_rec_maal.csv") |>
  mutate(assessment = "Current") |>
  dplyr::select(year, assessment, A0:A65) |>
  tidyr::pivot_longer(cols = c(-year, -assessment), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("A", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age),
  )

# Read old data
inputs_old <- r4ss::SS_read(dir = file.path("../../model/2017_yelloweye_model_updated_ss3_exe"))
wa_rec_maal_old <- inputs_old$dat$agecomp |>
  dplyr::filter(fleet == -7) |>
  mutate(assessment = "Previous") |>
  dplyr::select(year, assessment, Nsamp, `a0`:`a65`) |>
  tidyr::pivot_longer(cols = c(-year, -assessment, -Nsamp), names_to = "age", values_to = "freq") |>
  dplyr::mutate(age = gsub("^a", "", age)) |>
  dplyr::group_by(year) |>
  dplyr::mutate(
    freq = freq / sum(freq),
    age = as.numeric(age)
  )

  # Put the long data frames together and bubble plot!
together <- rbind(wa_rec_maal_old, wa_rec_maal_new)
comparison_plot_wa <- together |>
  dplyr::filter(freq > 0) |>
  ggplot2::ggplot(aes(x = year, y = age, col = assessment, size = freq)) +
  ggplot2::geom_point(position = position_dodge(0.5)) +
  ggtitle("WA_REC MAAL Comparison")
comparison_plot_wa
  
  #Save plot
ggsave(plot = comparison_plot_wa, "wa_rec_maal_comparisons.png", path = file.path("../../Rcode/length_age_comps"))
```