---
title: "rec_ages_for_datafile"
output: html_document
date: "2025-03-30"
---

Use this document as a source for Rec age comp code for pulling data into each of the model runs. I will comment out the code so that it is easy to see what should go into the models.

```{r setup, include=FALSE}
rm(list = ls())

library(dplyr)
library(tidyr)
library(ggplot2)
library(r4ss)
library(readr)
library(nwfscSurvey)
library(pacfintools)
```

Read in 2017 model data
```{r}
# Get inputs from 2017 assessment that will get carried over to this assessment
model_2017_path <- file.path("../../model/2017_yelloweye_model_updated_ss3_exe")
inputs <- SS_read(dir = model_2017_path, ss_new = TRUE)
```

1. CA_REC CAAL 1983-1996, aged by WDFW
  *no change* pull in old data and use in all model runs
  Future assessments: find the Traci file and make sure the ages and lengths are correct.

```{r}
ca_rec_caal_1983_1996 <- inputs$dat$agecomp |>
  filter(fleet == 3) |>
  filter(year == 1983 | year == 1996) |>
  filter(ageerr == 1)
```

2. CA_REC MAAL Ghost 1983-1996, aged by WDFW
  *no change* pull in old data and use in all model runs

```{r}
ca_rec_maal_1983_1996 <- inputs$dat$agecomp |>
  filter(fleet == -3) |>
  filter(year == 1983 | year == 1996) |>
  filter(ageerr == 1)
```

3. CA_REC CAAL Don Pearson Data 1979-1984, aged by Betty
  *divide all Nsamp and Ages by 2* - everything was doubled by accident in last assessment
```{r}
# Use this first section to grab the old data with out any changes! 
ca_rec_caal_Don <- inputs$dat$agecomp |>
  filter(fleet == 3) |>
  filter(year >= 1979 & year <= 1984) |>
  filter(ageerr == 2)

## For the Update, add this next section to correct for the doubling mistake
ca_rec_caal_Don[,9:75] <- ca_rec_caal_Don[,9:75]/2
```

4. CA_REC CAAL John Budrick Data 2009-2016, aged by Betty
  *divide all Nsamp and Ages by 2* - No new data, all data we got matched perfectly with 2017 assessment data, so just use old data.
```{r}
# Use this first section to grab the old data with out any changes! 
ca_rec_caal_John <- inputs$dat$agecomp |>
  filter(fleet == 3) |>
  filter(year >= 2009 & year <= 2016)

## For the Update, add this next section to correct for the doubling mistake
ca_rec_caal_John[,9:75] <- ca_rec_caal_John[,9:75]/2
```

5. CA_REC MAAL Ghost, Don's data
  *need to rebuild from CAAL* - Nsamp column was duplicated (and is wrong), so it shifted all of the ages forward by 1 and dropped the last age column
```{r}
# Use this first section to grab the old data with out any changes! 
old_ca_rec_maal_Don <- inputs$dat$agecomp |>
  filter(fleet == -3) |>
  filter(year >= 1979 & year <= 1984) |>
  filter(ageerr == 2)

## For the Update, add this next section to fix old mistake
# Take CAAL and group it so it matches MAAL structure, re-add the correct columns
ca_rec_maal_Don <- ca_rec_caal_Don %>%
  group_by(year) %>%  # Retain key columns
  summarise(across(starts_with("a"), \(x) sum(x, na.rm = TRUE)),  # Sum age columns
            Nsamp = sum(Nsamp, na.rm = TRUE), .groups = "drop")  # Sum input_n
ca_rec_maal_Don <- cbind(old_ca_rec_maal_Don[,1:8],ca_rec_maal_Don[,69],ca_rec_maal_Don[,3:68])
```

6. CA_REC MAAL Ghost, John's data
  *need to rebuild from CAAL* - Nsamp column is totally wrong, All the age data looks correct, but how they added the Nsamps together is off. Do the same thing as above and just rebuild from CAAL. Confirmed that the data matched the MAAL that I calculated using nwfsc code.
  Also, Vlada says that Nsamp column for MAAL data DOES NOT MATTER...but for an update, lets at least make it consistent with everything else in that Nsamp = "total samples".
  It is a lot easier to look for mistakes in the data if the Nsamps = total samples, and is consistent throughout the datafile.
```{r}
# Use this first section to grab the old data with out any changes! 
old_ca_rec_maal_John <- inputs$dat$agecomp |>
  filter(fleet == -3) |>
  filter(year >= 2009 & year <= 2016)

## For the Update, add this next section to fix old mistake
# Take CAAL and group it so it matches MAAL structure, re-add the correct columns
ca_rec_maal_John <- ca_rec_caal_John %>%
  group_by(year) %>%  # Retain key columns
  summarise(across(starts_with("a"), \(x) sum(x, na.rm = TRUE)),  # Sum age columns
            Nsamp = sum(Nsamp, na.rm = TRUE), .groups = "drop")  # Sum input_n
ca_rec_maal_John <- cbind(old_ca_rec_maal_John[,1:8],ca_rec_maal_John[,69],ca_rec_maal_John[,3:68])
```




7. OR_REC CAAL
  *doubling issue here too* - Also fixing Ageing Error column and adding updated Ages
  - If Ali or Vlada thinks we should ignore the new ages in 2015, then just use all old data and divide by 2
```{r}
# Use this first section to grab the old data with out any changes! 
# Bring in old data from 1979-2016
or_rec_caal <- inputs$dat$agecomp |>
  filter(fleet == 6) |>
  filter(year >= 1979)

# Next, fix doubling problem, without adding any new data
# Use for a model run where we fix issues before adding new data
or_rec_caal[,9:75] <- or_rec_caal[,9:75]/2

# Next problem is to fix the age column.Below is the correct age location and years according to Ali
# 1979 - 2000 = WDFW
# 2001 = WDFW(40) /unknown (assumed NWFSC) (10)
# 2002 =WDFW (n = 73)
# 2009 - 2016 = NWFSC
or_rec_caal[146:160,6] <- 1 # changing 2002 = WDFW
or_rec_caal[161:193,6] <- 2 # changing 2009-2016 = NWFSC

#################### Hold off on adding new data until Ali

# Last, Bring in new 2009-2016 data - some of the ages were updated (Check with Ali)
new_or_rec_caal <- read_csv(file.path("../../Data/processed/rec_comps/or_rec_caal.csv"))
names(new_or_rec_caal) <- names(or_rec_caal)
or_rec_caal <- rbind(or_rec_caal,new_or_rec_caal)
```

8. OR_REC MAAL ghost
  *why are Nsamps not whole numbers?* - It doesn't matter...but it is a lot easier to double check the data if Nsamps are "total samples", so rebuild using up to date CAAL data
```{r}
# Use this first section to grab the old data with out any changes!
# load old MAAL so we can use the first 8 columns and double check the data
old_or_rec_maal <- inputs$dat$agecomp |>
  filter(fleet == -6)

## For the Update, add this next section to fix Nsamps
# Take CAAL and group it so it matches MAAL structure, re-add the correct columns
or_rec_maal <- or_rec_caal %>%
  group_by(year) %>%  # Retain key columns
  summarise(across(starts_with("a"), \(x) sum(x, na.rm = TRUE)),  # Sum age columns
            Nsamp = sum(Nsamp, na.rm = TRUE), .groups = "drop")  # Sum input_n
or_rec_maal <- cbind(old_or_rec_maal[,1:8],or_rec_maal[,69],or_rec_maal[,3:68])
```

9. WA_REC CAAL
  *Doubling problem, however, Fabio (state rep) provided new data for the entire time series and recommends we use all new data* - some of the old ages have been updated.
```{r}
# Use this first section to grab the old data with out any changes!
old_wa_rec_caal <- inputs$dat$agecomp |>
  filter(fleet == 7)

# Use this section to fix the doubling problem for OLD data
old_wa_rec_caal[,9:75] <- old_wa_rec_caal[,9:75]/2

## For the Update, add this next section to use the updated ages (Fabio), ignore all old data
#load new processed data from most recent RecFin pull
wa_rec_caal <- read_csv(file.path("../../Data/processed/rec_comps/wa_rec_caal.csv"))
names(wa_rec_caal) <- names(old_wa_rec_caal) # Replace col names from another MAAL dataset so they match everything else

```

10. WA_REC MAAL
  *Just use new data* - some of the old ages have been updated.
```{r}
# Use this first section to grab the old data with out any changes!
old_wa_rec_maal <- inputs$dat$agecomp |>
  filter(fleet == -7)

## For the Update, add this next section to use the updated ages (Fabio), ignore all old data
#load new processed data from most recent RecFin pull
wa_rec_maal <- read_csv(file.path("../../Data/processed/rec_comps/wa_rec_maal.csv"))
names(wa_rec_maal) <- names(old_wa_rec_maal) # Replace col names from another MAAL dataset so they match everything else
```

Finish and put it all together 
```{r}
# Always double check the names we are calling, they may change depending on if we use the old vs the new data
mod$dat$agecomp <- dplyr::bind_rows(ca_rec_caal_1983_1996,
                                  ca_rec_maal_1983_1996, 
                                  ca_rec_caal_Don, 
                                  ca_rec_caal_John, 
                                  ca_rec_maal_Don,
                                  ca_rec_maal_John,
                                  or_rec_caal,
                                  or_rec_maal,
                                  wa_rec_caal,
                                  wa_rec_maal) 
# list all the dfs that I just made, when we start to add others, make sure to insert Commercial comps in the correct spot
```
