---
title: "rec_ages"
output: html_document
date: "2025-03-13"
---

```{r setup}
library(dplyr)
library(tidyr)
library(ggplot2)
library(r4ss)
library(readr)
library(nwfscSurvey)
library(pacfintools)
```

*if there are no new CA age data since 2016, then dont worry about it*

Load in all data sources
1. ??? where is all the data for this

Fleets:
3 - CA_REC
6 - OR_REC 
7 - WA_REC (abby)

```{r}
#recfin_bio <- read_csv("C:/Morgan's Stuff/Sebastes_ruberrimus_2025/Data/raw/confidential/SD501.csv")

```

#Data processing: select and filter relevant data, then organize to match format of SS file. Then, load 2017 data and compare with bubble plots.


CA_REC: CAAL, CDFW. We only need to find CDFW data starting in 2009 (John Budrick)

```{r}

```

CA_REC: MAAL, CDFW. We only need to find CDFW data starting in 2009 (John Budrick)

```{r}

```

OR_REC: CAAL, WDFW ager data. Starting in 1979

```{r}

```

OR_REC: MAAL, WDFW ager data. Starting in 1979

```{r}

```

WA_REC: CAAL, WDFW ager data. Starting in 1998. includes MRFSS and OSP

```{r}

```

WA_REC: MAAL, WDFW ager data. Starting in 1998. includes MRFSS and OSP

```{r}

```